---
title: "Physics Validation of Y3 Clusters Code"
author: "Y3 Cluster Team"
date: "`r Sys.Date()`"
output:
  html_document:
      toc: true
      toc_depth: 1
      number_sections: true
  pdf_document: default
  word_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE,
                      fig.align = "center",
                      fig.asp = 1/3)
library(tidyverse)
truecolor = "red"
trueshape = 1
testcolor = "blue"
testshape = 3
frac.errormargin = 1.0e-6
abs.errormargin = 1.0e-12
textfor <- function(bool) {
  checkmate::assert_logical(bool, len = 1)
  if (bool) return("passes")
  return("<span style=\"color:red\">fails</span>")
}
testresult <- function(df) textfor(all(df$pass))

geom_smallpoint <- function(...) geom_point(..., size = 0.5)

nunique <- function(x) length(unique(x))
```

# Introduction

This document presents
demonstrations of the validity of calculations
done in the Y3 clusters analysis code.
Each section contains the validation testing results
for a single function in our calculation.
For each of these functions,
we have results from a trusted sources that are assumed to be correct,
evaluated at some number of points.
We test our approximations
by comparing them to the expected answer at each of these points.
Our unit tests require fractional errors to be less than
$`r frac.errormargin`$,
**or** absolute errors less than
$`r abs.errormargin`$,
unless otherwise indicated.
For a test to pass,
it is required that every calculated point pass.

In each section, we show both absolute and residual comparisons.
The absolute comparisons are shown
to discover any gross discrepencies
between the correct result and our approximations.
In the absolute comparison plots,
the *expected* (or *true*) values are shown in `r truecolor`,
and our *approximations* (or *test* values) are shown in `r testcolor`.
The residual comparisons show more detail.
In each case, we show fractional (relative) residuals;
we plot the absolute residuals only when the fractional residual test fails
for one or more points.
Because the criterion for the calculation at each point to pass is a logical
**or** of the comparison to the relative and absolute error bounds,
each plot contains information about both tests.
In the plot of the fractional residuals,
the color of each point is determined by whether or not that point
passes the absolute residual test.
In the plot of the absolute residuals,
the color of each point is determined by whether or not that point
passes the fractional residual test.
Thus, in each plot, a point passes the test if *either* it is within
the error envelope shown by the blue horizontal lines on the plot,
*or* it is blue.
Points that fail the test will be both outside of the allowable error
envelope and red.

# $\Omega_z$

```{r}
local.frac.errormargin = frac.errormargin
local.abs.errormargin = abs.errormargin
```

```{r, include = FALSE}
omega_z <- read_tsv("../data/omega_z_sdss.out")
omega_z <- mutate(omega_z,
                  resid = ytrue - ytest,
                  fracerr = resid/ytrue,
                  lo = pmin(-local.frac.errormargin * ytrue, -local.abs.errormargin),
                  hi = pmax(local.frac.errormargin * ytrue, local.abs.errormargin),
                  within.frac = (abs(fracerr) < local.frac.errormargin),
                  within.abs = (abs(resid) < local.abs.errormargin),
                  pass = within.frac | within.abs)
```

$\Omega_z$ is the area of the survey as a function of redshift $z$.
Our calculation is implemented in the file `omega_z_sdss.hh`.
Our testing uses `r nunique(omega_z$z)` $z$ values.
Over the range tested
$(`r min(omega_z$z)` < z < `r max(omega_z$z)`)$,
the maximum magnitude of the absolute residual between the expected value
and our approximation is
$`r max(abs(omega_z$resid))`$;
the maximum magnitude of the fractional residual is
$`r max(abs(omega_z$fracerr))`$.
For this test, we require the relative residuals to be less than
$`r local.frac.errormargin`$,
or the absolute residuals to be less than
$`r local.abs.errormargin`$.

This test `r testresult(omega_z)`.

```{r fig.cap = "Absolute comparision of our approximation to the expected results for $\\Omega_z$."}
ggplot(omega_z) +
  geom_smallpoint(aes(x = z, y = ytrue), color = truecolor, shape = trueshape) +
  geom_smallpoint(aes(x = z, y = ytest), color = testcolor, shape = testshape) +
  xlab(expression(z)) +
  ylab(expression(Omega))
```

```{r fig.cap = "Relative residual comparison or our approximation to the expected results for $\\Omega_z$. Points in blue pass, and points in red fail, the absolute residual test."}
ggplot(omega_z, aes(x = z, y = fracerr, color = within.abs)) +
  geom_smallpoint() +
  guides(color = FALSE) +
  ylim(-2 * local.frac.errormargin, 2 * local.frac.errormargin) +
  xlab(expression(z)) +
  ylab("fractional residual") +
  geom_hline(yintercept = local.frac.errormargin, color="blue") +
  geom_hline(yintercept = -1 * local.frac.errormargin, color="blue")
```

# Mass-observable relation

```{r}
local.frac.errormargin = 5 * frac.errormargin
local.abs.errormargin = abs.errormargin
```

```{r include = FALSE}
mor_t2 <- read_tsv("../data/mor_tt_test.out")
mor_t2$mbin <- as.factor(mor_t2$log10m)
mor_t2 <- mutate(mor_t2,
                 resid = probtrue - probtest,
                 fracerr = resid/probtrue,
                 lo = pmin(-local.frac.errormargin * probtrue, -local.abs.errormargin),
                 hi = pmax(local.frac.errormargin * probtrue, local.abs.errormargin),
                 within.frac = (abs(fracerr) < local.frac.errormargin),
                 within.abs = (abs(resid) < local.abs.errormargin),
                 pass = within.frac | within.abs)
```

*We need a description of what probability is being calculated.*
Our calculation is implemented in the file `mor_t2.hh`.
Our testing uses `r nlevels(mor_t2$mbin)` different mass values,
and `r nunique(mor_t2$lambda)` different $\lambda$ values.
Over the range tested 
($`r min(mor_t2$lambda)` < \lambda < `r max(mor_t2$lambda)`$),
the maximum magnitude of the absoluteresidual between the expected value
and our calculated value is
$`r max(abs(mor_t2$resid))`$;
the maximum magnitude of the fractional residual is
$`r max(abs(mor_t2$fracerr))`$.
For this test,
we require the relative differences to be 
less than $`r local.frac.errormargin`$,
**or** the absolute difference to be less than
$`r local.abs.errormargin`$.

This test `r testresult(mor_t2)`.

Each panel in the figure below corresponds to a (ln) mass bin;
the value of the (ln) mass is specified in the panel top label.

```{r fig.cap = "Absolute comparsion of our approximation to the expected results for `mor_t2`."}
ggplot(mor_t2) +
  geom_smallpoint(aes(x = lambda, y = probtrue), color = truecolor, shape = trueshape) +
  geom_smallpoint(aes(x = lambda, y = probtest), color = testcolor, shape = testshape) +
  xlab(expression(lambda)) +
  ylab("probability") +
  facet_wrap(mbin~.)
```

```{r fig.cap = "Relative residual comparison of our approximation to the expected results for mor_t2. Blue points pass, and red points fail, the absolute residual test.", warning = FALSE}
ggplot(mor_t2, aes(x = lambda, y = fracerr, color = within.abs)) +
  geom_smallpoint() +
  guides(color = FALSE) +
  xlab(expression(lambda)) +
  ylab("fractional residual") +
  geom_hline(yintercept = local.frac.errormargin, color = "blue") +
  geom_hline(yintercept = -1 * local.frac.errormargin, color = "blue") +
  ylim(-2 * local.frac.errormargin, 2 * local.frac.errormargin) +
  facet_wrap(mbin ~ .)
```

```{r fig.cap = "Absolute residual comparison of our approximation to the expected results for mor_t2. Blue points pass, and red points fail, the relative residual test.", warning = FALSE}
ggplot(mor_t2, aes(x= lambda, y = resid, color = within.frac)) +
  geom_smallpoint() +
  guides(color = FALSE) +
  geom_hline(yintercept = local.abs.errormargin, color = "blue") +
  geom_hline(yintercept = -1 * local.abs.errormargin, color = "blue") +
  ylim(-2 * local.abs.errormargin, 2 * local.abs.errormargin) +
  xlab(expression(lambda)) +
  facet_wrap(mbin ~ .)
```

# Session information

The following R package versions were used to prepare this document.

```{r echo = TRUE}
sessionInfo()
```

