---
title: "Physics Validation of Y3 Clusters Code"
author: "Y3 Cluster Team"
date: "12/12/2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE,
                      fig.align = "center")
library(tidyverse)
```

## Introduction

This document presents analysis to demonstrate the calculations done by the Y3 clusters analysis code.

## An example

This is an example of how to read data files and make plots.
It is not an actually interesting example of a plot.
We can settle upon standard formatting, etc., at a later date.

### Absolute comparison of 1-d interpolation results to exact answers

```{r, include = FALSE}
d <- read_tsv("../data/interp_1d.out")
d <- mutate(d, resid = ytrue - ytest)
```

```{r, fig.cap = "Comparison of linear interpolation and exact results"}
ggplot(d) +
  geom_point(aes(x, ytrue), color = "red") +
  geom_point(aes(x, ytest), color = "blue") +
  xlab("x") +
  ylab("y")
```

### Fractional comparison of linear interpolation and exact results

```{r, fig.cap = "Difference between linear interpolation and exact results"}
ggplot(d) +
  geom_point(aes(x, resid/ytrue)) +
  xlab("x") +
  ylab("residual/ytrue")
```
