---
title: "Physics Validation of Y3 Clusters Code"
author: "Y3 Cluster Team"
date: "12/12/2018"
output:
  html_document: default
  pdf_document: default
  word_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE,
                      fig.align = "center"
                      )
library(tidyverse)
truecolor = "red"
trueshape = 1
testcolor = "blue"
testshape = 3
testresult <- function(bool) {
  checkmate::assert_logical(bool, len = 1)
  if (bool) return("passes")
  return("<span style=\"color:red\">fails</span>")
}
```

## Introduction

This document presents analysis to demonstrate the calculations done by the Y3 clusters analysis code.

Our unit tests require fractional errors to be less than one part per million,
unless otherwise indicated.

In plots that show both *true* and *test* values,
the *true* value plotted in `r truecolor`
and the *test* value is plotted in `r testcolor`.

## Omega_z

$\Omega_z$ is the area of the survey as a function of redshift $z$.
Our calculation is implemented in the file `omega_z_sdss.hh`.

```{r, include = FALSE}
omega_z <- read_tsv("../data/omega_z_sdss.out")
omega_z <- mutate(omega_z, resid = ytrue - ytest)
```

### Absolute comparison of omega_z

```{r fig.cap = "Absolute comparision", fig.asp = 1/2}
ggplot(omega_z) +
  geom_point(aes(x = z, y = ytrue), color = truecolor, shape = trueshape) +
  geom_point(aes(x = z, y = ytest), color = testcolor, shape = testshape) +
  xlab("redshift") +
  ylab("survey area")  
```

### Relative error

Over the range tested (`r min(omega_z$z)` < $z$ < `r max(omega_z$z)`),
the maximum magnitude of deviation between the expected value
and our calculated value is
$`r max(abs(omega_z$resid))`$;
the maximum absolute fractional deviation is
$`r max(abs(omega_z$resid/omega_z$ytrue))`$.

For this test, we require the differences to be less than $10^{-6}$.
This test `r testresult((max(abs(omega_z$resid)) < 1.0e-6))`. 

```{r fig.cap = "Relative comparison"}
ggplot(omega_z, aes(x = z, y = resid/ytrue)) +
  geom_point() +
  ylim(-1.0e-6, 1.0e-6) +
  xlab("redshift") +
  ylab("relative difference in survey area") 
```

## Session information

The following R package versions were used to prepare this document.

```{r echo = TRUE}
sessionInfo()
```

