---
title: "Physics Validation of Y3 Clusters Code"
author: "Y3 Cluster Team"
date: "12/12/2018"
output:
  html_document: default
  pdf_document: default
  word_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE,
                      fig.align = "center"
                      )
library(tidyverse)
truecolor = "red"
trueshape = 1
testcolor = "blue"
testshape = 3
errormargin = 1.0e-6
testresult <- function(bool) {
  checkmate::assert_logical(bool, len = 1)
  if (bool) return("passes")
  return("<span style=\"color:red\">fails</span>")
}
```

## Introduction

This document presents analysis to demonstrate the calculations done by the Y3 clusters analysis code.

Our unit tests require fractional errors to be less than `r errormargin`,
unless otherwise indicated.

In plots that show both *true* and *test* values,
the *true* value plotted in `r truecolor`
and the *test* value is plotted in `r testcolor`.

## Omega_z

$\Omega_z$ is the area of the survey as a function of redshift $z$.
Our calculation is implemented in the file `omega_z_sdss.hh`.

```{r, include = FALSE}
omega_z <- read_tsv("../data/omega_z_sdss.out")
omega_z <- mutate(omega_z, resid = ytrue - ytest)
```

### Absolute comparison of omega_z

```{r fig.cap = "Absolute comparision", fig.asp = 1/2}
ggplot(omega_z) +
  geom_point(aes(x = z, y = ytrue), color = truecolor, shape = trueshape) +
  geom_point(aes(x = z, y = ytest), color = testcolor, shape = testshape) +
  xlab("redshift") +
  ylab("survey area")  
```

### Relative error

Over the range tested (`r min(omega_z$z)` < $z$ < `r max(omega_z$z)`),
the maximum magnitude of deviation between the expected value
and our calculated value is
$`r max(abs(omega_z$resid))`$;
the maximum absolute fractional deviation is
$`r max(abs(omega_z$resid/omega_z$ytrue))`$.

For this test, we require the differences to be less than $`r errormargin`$.
This test `r testresult((max(abs(omega_z$resid)) < errormargin))`.
The red lines indicate the allowable error bounds.

```{r fig.cap = "Relative comparison", fig.asp = 1/2}
ggplot(omega_z, aes(x = z, y = resid/ytrue)) +
  geom_point() +
  ylim(-1 * errormargin, errormargin) +
  xlab("redshift") +
  ylab("relative difference in survey area") +
  geom_hline(yintercept = errormargin, color="red") +
  geom_hline(yintercept = -1 * errormargin, color="red")
```

## Mass observable relation (MOR)

```{r include = FALSE}
mor_t2 <- read_tsv("../data/mor_tt_test.out")
mor_t2$mbin <- as.factor(mor_t2$log10m)
mor_t2 <- mutate(mor_t2, resid = probtrue - probtest)
```

Our calculation is implemented in the file `mor_t2.hh`.
Our testing uses `r nlevels(mor_t2$mbin)` different mass values,
and many $\lambda$ values.

### Absolute comparision

Each panel in the figure below corresponds to a (log) mass bin;
the value of the (log) mass is specified in the panel top label.

```{r fig.cap = "Absolute comparsion", fig.asp = 1/2}
ggplot(mor_t2) +
  geom_point(aes(x = lambda, y = probtrue), color = truecolor, shape = trueshape) +
  geom_point(aes(x = lambda, y = probtest), color = testcolor, shape = testshape) +
  xlab(expression(lambda)) +
  ylab("probability") +
  facet_wrap(mbin~.)
```

### Relative comparison

Over the range tested (`r min(mor_t2$lambda)` < $\lambda$ < `r max(mor_t2$lambda)`),
the maximum magnitude of deviation between the expected value
and our calculated value is
$`r max(abs(mor_t2$resid))`$;
the maximum absolute fractional deviation is
$`r max(abs(mor_t2$resid/mor_t2$probtrue))`$.

For this test, we require the differences to be less than $`r errormargin`$.
This test `r testresult((max(abs(mor_t2$resid/mor_t2$probtrue)) < errormargin))`.
The red lines indicate the allowable error bounds.

```{r fig.cap = "Relative comparison", fig.asp = 1/2}
ggplot(mor_t2, aes(x = lambda, y = resid/probtrue)) +
  geom_point() +
  xlab(expression(lambda)) +
  ylab("relative difference probability") +
  geom_hline(yintercept = errormargin, color="red") +
  geom_hline(yintercept = -1 * errormargin, color="red") +
  facet_wrap(mbin ~ .)
```

## Session information

The following R package versions were used to prepare this document.

```{r echo = TRUE}
sessionInfo()
```

