---
title: "Physics Validation of Y3 Clusters Code"
author: "Y3 Cluster Team (3ct)"
date: "`r Sys.Date()`"
output:
  html_document:
      toc: true
      toc_depth: 1
      toc_float: true
      number_sections: true
  pdf_document: default
  word_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE,
                      cache = FALSE,
                      fig.align = "center",
                      fig.asp = 1/3)
library(tidyverse)
truecolor = "black"
trueshape = 1
testcolor = "cyan"
testshape = 3
frac.errormargin = 1.0e-6
abs.errormargin = 1.0e-12

testresult <-function(df) {
  npass <- sum(df$pass)
  npoints <- nrow(df)
  str_interp("**For this test, ${sum(df$pass)} of ${nrow(df)} points pass.**")
}

geom_smallpoint <- function(...) geom_point(..., size = 0.5)

nunique <- function(x) length(unique(x))
```

# Introduction

This document presents
demonstrations of the validity of calculations
done in the Y3 clusters analysis code.
Each section contains the validation testing results
for a single function in our calculation.

a. For each of these functions,
we have evaluations from a trusted source that are assumed to be correct.
We test *our implementation* by comparing the results to the evaluations from the trusted source.
a. Unless otherwise indicated, we require fractional differences to be less than
$`r frac.errormargin`$,
**or** absolute differences less than
$`r abs.errormargin`$.
The relative error bound was chosen to be
smaller than the systematic uncertainty
to which the value is known from the trusted source.
An absolute error bound is needed for the cases in which
the predicted value is 0 or very close to 0.
a. In each section, we show both absolute and residual comparisons.
In the absolute comparison plots,
the *expected* (or *true*) values are shown in `r truecolor`,
and results from our *implementation* (or *test* values) are shown in `r testcolor`. We also show the fractional (relative) and absolute residuals.
a. In the plot of the fractional residuals,
the color of each point is determined by whether or not that point
passes the absolute residual test.
In the plot of the absolute residuals,
the color of each point is determined by whether or not that point
passes the fractional residual test.
Thus, in each plot, a point passes the test if *either* it is blue or within
the error bounds (indicated by horizontal lines).
Points that fail the test will be both red and outside of the allowable error bounds (indicated by horizontal lines).
a. For a test to pass, it is required that every calculated point pass.
This condition is verified automatically and the result is noted in bold in each section.


# Survey Area

## SDSS Survey footprint function

```{r}
local.frac.errormargin = frac.errormargin
local.abs.errormargin = abs.errormargin
```

```{r, include = FALSE}
omega_z <- read_tsv("../release-build/data/omega_z_sdss.out")
omega_z <- mutate(omega_z,
                  resid = ytrue - ytest,
                  fracerr = resid/ytrue,
                  lo = pmin(-local.frac.errormargin * ytrue, -local.abs.errormargin),
                  hi = pmax(local.frac.errormargin * ytrue, local.abs.errormargin),
                  within.frac = (abs(fracerr) < local.frac.errormargin),
                  within.abs = (abs(resid) < local.abs.errormargin),
                  pass = within.frac | within.abs)
```

$\Omega_z$ is the area of the survey as a function of redshift $z$.
This section tests our implementation of the SDSS survey area function, which is implemented in the file `omega_z_sdss.hh`.

For this testing, we compare our function output to the independent function evaluations from Matteo Costanzi in the SDSS cluster cosmology analysis (Costanzi et al. 2019).

a. The comparisons are done at `r nrow(omega_z)` unique $z$ points. 
a. Over the range tested
$(`r min(omega_z$z)` < z < `r max(omega_z$z)`)$,
the maximum magnitude of the absolute residual between the expected value
and our approximation is
$`r max(abs(omega_z$resid))`$;
the maximum magnitude of the fractional residual is
$`r max(abs(omega_z$fracerr))`$.
a. For this test, we require the relative residuals to be less than
$`r local.frac.errormargin`$,
or the absolute residuals to be less than
$`r local.abs.errormargin`$.
a. `r testresult(omega_z)`

```{r fig.cap = "Absolute comparison of our implementation results (cyan points) to the expected values (black points) for $\\Omega_z$. Note that over-plotting may obscure the expected values."}
ggplot(omega_z) +
  geom_smallpoint(aes(x = z, y = ytrue), color = truecolor, shape = trueshape) +
  geom_smallpoint(aes(x = z, y = ytest), color = testcolor, shape = testshape) +
  xlab(expression(z)) +
  ylab(expression(Omega))
```

```{r fig.cap = "Relative residual between our implementation results and the expected values. Blue horizontal lines are the error tolerance bounds. Blue points pass, and red points fail, the absolute residual test."}
ggplot(omega_z, aes(x = z, y = fracerr, color = within.abs)) +
  geom_smallpoint() +
  guides(color = "none") +
  ylim(-2 * local.frac.errormargin, 2 * local.frac.errormargin) +
  xlab(expression(z)) +
  ylab("fractional residual") +
  geom_hline(yintercept = local.frac.errormargin, color="blue") +
  geom_hline(yintercept = -1 * local.frac.errormargin, color="blue")
```

# Halo Mass Function

**This test requires running masses_test.ini in cosmosis_tests first**
Need to do better than this in the future.

```{r}
local.frac.errormargin = 0.01
local.abs.errormargin = 1.0e-16
```

```{r include = FALSE}
hmf <- read_tsv("../data/masses_test.out")
hmf$zbin <- as.factor(hmf$z)
hmf <- mutate(hmf,
                 resid = ytrue - ytest,
                 fracerr = resid/ytrue,
                 lo = pmin(-local.frac.errormargin * ytrue, -local.abs.errormargin),
                 hi = pmax(local.frac.errormargin * ytrue, local.abs.errormargin),
                 within.frac = (abs(fracerr) < local.frac.errormargin),
                 within.abs = (abs(resid) < local.abs.errormargin),
                 pass = within.frac | within.abs)
```

$n(M_{200m}, z)$ is the halo mass function, which describes the 3D cosmic density of halos at a given mass $M_200m$ and $z$. 
We have modified the CosmoSIS halo mass function module to consider the effect of massive neutrinos (Costanzi et al. 2013).

For this testing, we compare results from our implementation to the  independent function evaluations from Matteo Costanzi in the SDSS cluster cosmology analysis (Costanzi et al. 2019)

a. The comparisons are made at `r nunique(hmf$z)` unique $z$ values with `r nlevels(hmf$zbin)` different mass values.
a. Over the range tested 
($`r min(hmf$mass)` < M < `r max(hmf$mass)`$),
the maximum magnitude of the absolute residual between the expected value
and our calculated value is
$`r max(abs(hmf$resid))`$;
the maximum magnitude of the fractional residual is
$`r max(abs(hmf$fracerr))`$.
a. For this test,
we require the relative differences to be 
less than $`r local.frac.errormargin`$,
or the absolute difference to be less than
$`r local.abs.errormargin`$.
a. `r testresult(hmf)`

Each panel in the figure below corresponds to a different $z$ value, specified in the panel top label.

```{r fig.asp=0.4,fig.cap = "Absolute comparison of our implementation to the expected results. Note that over-plotting may obscure the expected values, if the approximated values are very close."}
ggplot(hmf) +
  geom_smallpoint(aes(x = mass, y = ytrue), color = truecolor, shape = trueshape) +
  geom_smallpoint(aes(x = mass, y = ytest), color = testcolor, shape = testshape) +
  xlab(expression(mass)) +
  ylab("Number Density") +
  scale_y_log10() +
  facet_wrap(zbin~ .)
```

```{r fig.asp=0.4,fig.cap = "Relative residual comparison of our implementation to the expected results. Blue points pass, and red points fail, the absolute residual test.", warning = FALSE}
ggplot(hmf, aes(x = mass, y = fracerr, color = within.abs)) +
  geom_smallpoint() +
  guides(color = "none") +
  xlab(expression(M)) +
  ylab("fractional residual") +
  geom_hline(yintercept = local.frac.errormargin, color = "blue") +
  geom_hline(yintercept = -1 * local.frac.errormargin, color = "blue") +
  ylim(-2 * local.frac.errormargin, 2 * local.frac.errormargin) +
  facet_wrap(zbin~ .)
```

# Volume Element

```{r}
local.frac.errormargin = 1.0e-3
local.abs.errormargin = abs.errormargin
```

```{r, include = FALSE}
dvdodz <- read_tsv("../release-build/data/dvdodz_test.out")
dvdodz <- mutate(dvdodz,
                  resid = ytrue - ytest,
                  fracerr = resid/ytrue,
                  lo = pmin(-local.frac.errormargin * ytrue, -local.abs.errormargin),
                  hi = pmax(local.frac.errormargin * ytrue, local.abs.errormargin),
                  within.frac = (abs(fracerr) < local.frac.errormargin),
                  within.abs = (abs(resid) < local.abs.errormargin),
                  pass = within.frac | within.abs)
```

$\frac{dV}{d\Omega dz}$ is the co-moving volume element in solid angle $d \Omega$ and redshift interval $dz$, and is a function of $z$. Our implementation is in the file `dv_do_dz_t.hh`.

For this testing, we compare our function output to the independent function evaluations from Matteo Costanzi in the SDSS cluster cosmology analysis (Costanzi et al. 2019), calculated with $\Omega_m =0.3$,  $\Omega_\Lambda =0.7$, $h_0=0.7$.

a. The comparisons are done at `r nrow(dvdodz)` unique $z$ values.
a. Over the range tested
$(`r min(dvdodz$z)` < z < `r max(dvdodz$z)`)$,
the maximum magnitude of the absolute residual between the expected value
and our approximation is
$`r max(abs(dvdodz$resid))`$;
the maximum magnitude of the fractional residual is
$`r max(abs(dvdodz$fracerr))`$.
a. For this test, we require the relative residuals to be less than
$`r local.frac.errormargin`$,
or the absolute residuals to be less than
$`r local.abs.errormargin`$.
a. `r testresult(dvdodz)`

```{r fig.cap = "Absolute comparison of our approximation to the expected results. Note that over-plotting may obscure the expected values, if the approximated values are very close."}
ggplot(dvdodz) +
  geom_smallpoint(aes(x = z, y = ytrue), color = truecolor, shape = trueshape) +
  geom_smallpoint(aes(x = z, y = ytest), color = testcolor, shape = testshape) +
  xlab(expression(z)) +
  ylab(expression(dvdodz))
```

```{r fig.cap = "Relative residual comparison or implementation to the expected results. Points in blue pass, and points in red fail, the absolute residual test."}
ggplot(dvdodz, aes(x = z, y = fracerr, color = within.abs)) +
  geom_smallpoint() +
  guides(color = "none") +
  ylim(-local.frac.errormargin, local.frac.errormargin) +
  xlab(expression(z)) +
  ylab("fractional residual") +
  geom_hline(yintercept = -local.frac.errormargin, color="blue") +
  geom_hline(yintercept =local.frac.errormargin, color="blue")
```

# Scaling Relations

## Mass-observable relation

```{r}
local.frac.errormargin = 5 * frac.errormargin
local.abs.errormargin = abs.errormargin
```

```{r include = FALSE}
mor_t2 <- read_tsv("../release-build/data/mor_tt_test.out")
mor_t2$mbin <- as.factor(mor_t2$log10m/log(10))
mor_t2 <- mutate(mor_t2,
                 resid = probtrue - probtest,
                 fracerr = resid/probtrue,
                 lo = pmin(-local.frac.errormargin * probtrue, -local.abs.errormargin),
                 hi = pmax(local.frac.errormargin * probtrue, local.abs.errormargin),
                 within.frac = (abs(fracerr) < local.frac.errormargin),
                 within.abs = (abs(resid) < local.abs.errormargin),
                 pass = within.frac | within.abs)
```

$P(\lambda |M_{200m})$ describes the probabilistic distribution of the cluster observable richness ($\lambda$) at a given mass $M_200m$. 
Our calculation is implemented in the file `mor_t2.hh`.

For this testing, we compare results from our implementation to the  independent function evaluations from Matteo Costanzi in the SDSS cluster cosmology analysis (Costanzi et al. 2019)

a. The comparisons are made at `r nunique(mor_t2$lambda)` unique $\lambda$ values with `r nlevels(mor_t2$mbin)` different mass values.
a. Over the range tested 
($`r min(mor_t2$lambda)` < \lambda < `r max(mor_t2$lambda)`$),
the maximum magnitude of the absolute residual between the expected value
and our calculated value is
$`r max(abs(mor_t2$resid))`$;
the maximum magnitude of the fractional residual is
$`r max(abs(mor_t2$fracerr))`$.
a. For this test,
we require the relative differences to be 
less than $`r local.frac.errormargin`$,
or the absolute difference to be less than
$`r local.abs.errormargin`$.
a. `r testresult(mor_t2)`

Each panel in the figure below corresponds to a different $\mathrm{log} M_{200m}$ value, specified in the panel top label.

```{r fig.asp=0.4,fig.cap = "Absolute comparison of our implementation to the expected results. Note that over-plotting may obscure the expected values, if the approximated values are very close."}
ggplot(mor_t2) +
  geom_smallpoint(aes(x = lambda, y = probtrue), color = truecolor, shape = trueshape) +
  geom_smallpoint(aes(x = lambda, y = probtest), color = testcolor, shape = testshape) +
  xlab(expression(lambda)) +
  ylab("probability") +
  facet_wrap(mbin~.)
```

```{r fig.asp=0.4,fig.cap = "Relative residual comparison of our implementation to the expected results. Blue points pass, and red points fail, the absolute residual test.", warning = FALSE}
ggplot(mor_t2, aes(x = lambda, y = fracerr, color = within.abs)) +
  geom_smallpoint() +
  guides(color = "none") +
  xlab(expression(lambda)) +
  ylab("fractional residual") +
  geom_hline(yintercept = local.frac.errormargin, color = "blue") +
  geom_hline(yintercept = -1 * local.frac.errormargin, color = "blue") +
  ylim(-2 * local.frac.errormargin, 2 * local.frac.errormargin) +
  facet_wrap(mbin~ .)
```

```{r fig.asp=0.4, fig.cap = "Absolute residual comparison of our approximation to the expected results for MOR. Blue points pass, and red points fail, the relative residual test.", warning = FALSE}
ggplot(mor_t2, aes(x= lambda, y = resid, color = within.frac)) +
  geom_smallpoint() +
  guides(color = "none") +
  geom_hline(yintercept = local.abs.errormargin, color = "blue") +
  geom_hline(yintercept = -1 * local.abs.errormargin, color = "blue") +
  ylim(-2 * local.abs.errormargin, 2 * local.abs.errormargin) +
  xlab(expression(lambda)) +
  facet_wrap(mbin ~ .)
```

## Projection effect on Scaling relations

```{r}
local.frac.errormargin =  frac.errormargin
local.abs.errormargin = abs.errormargin
```

```{r include = FALSE}
lc_lt <- read_tsv("../release-build/data/lc_lt_test.out")
lc_lt$lbin <- as.factor(lc_lt$lambda_true)
lc_lt <- mutate(lc_lt,
                 resid = probtrue - probtest,
                 fracerr = resid/probtrue,
                 lo = pmin(-local.frac.errormargin * probtrue, -local.abs.errormargin),
                 hi = pmax(local.frac.errormargin * probtrue, local.abs.errormargin),
                 within.frac = (abs(fracerr) < local.frac.errormargin),
                 within.abs = (abs(resid) < local.abs.errormargin),
                 pass = within.frac | within.abs)
```

Projections of LOS structures cause a shift and an extra scatter to the values of cluster richness $\lambda$. $P(\lambda_c |\lambda_t)$ describes the probabilistic distribution of the richness scatter (between $\lambda_c$ and $\lambda_t$) caused by projections. 
Our calculation is implemented in the file `lc_lt_t.hh`.

For this testing, we compare results from our implementation to the  independent function evaluations from Matteo Costanzi's projection analyses (Costanzi et al. 2019)

a. The comparisons are made at `r nunique(lc_lt$lambda_cen)` unique $\lambda_{c}$ values with `r nlevels(lc_lt$lbin)` different $\lambda_{t}$ values.
a. Over the range tested 
($`r min(lc_lt$lambda_cen)` < \lambda_{cen} < `r max(lc_lt$lambda_cen)`$),
the maximum magnitude of the absolute residual between the expected value
and our calculated value is
$`r max(abs(lc_lt$resid))`$;
the maximum magnitude of the fractional residual is
$`r max(abs(lc_lt$fracerr))`$.
a. For this test,
we require the relative differences to be 
less than $`r local.frac.errormargin`$,
or the absolute difference to be less than
 $`r local.abs.errormargin`$.
a. `r testresult(lc_lt)`

Each panel in the figure below corresponds to a $\lambda_{t}$ value,  specified in the panel top label.

```{r fig.asp=0.7, fig.cap = "Absolute comparison of our implementation to the expected results. Note that over-plotting may obscure the expected values, if the approximated values are very close."}
ggplot(lc_lt) +
  geom_smallpoint(aes(x = lambda_cen, y = probtrue), color = truecolor, shape = trueshape) +
  geom_smallpoint(aes(x = lambda_cen, y = probtest), color = testcolor, shape = testshape) +
  xlab(expression(lambda_c)) +
  ylab("probability") +
  facet_wrap(lbin~.)
```

```{r fig.asp=0.7, fig.cap = "Relative residual comparison of our implementation to the expected results. Blue points pass, and red points fail, the absolute residual test.", warning = FALSE}
ggplot(lc_lt, aes(x = lambda_cen, y = fracerr, color = within.abs)) +
  geom_smallpoint() +
  guides(color = "none") +
  xlab(expression(lambda_c)) +
  ylab("fractional residual") +
  geom_hline(yintercept = local.frac.errormargin, color = "blue") +
  geom_hline(yintercept = -1 * local.frac.errormargin, color = "blue") +
  ylim(-2 * local.frac.errormargin, 2 * local.frac.errormargin) +
  facet_wrap(lbin~ .)
```

```{r fig.asp=0.7, fig.cap = "Absolute residual comparison of our implementation to the expected results. Blue points pass, and red points fail, the relative residual test.", warning = FALSE}
ggplot(lc_lt, aes(x= lambda_cen, y = resid, color = within.frac)) +
  geom_smallpoint() +
  guides(color = "none") +
  geom_hline(yintercept = local.abs.errormargin, color = "blue") +
  geom_hline(yintercept = -1 * local.abs.errormargin, color = "blue") +
  ylim(-2 * local.abs.errormargin, 2 * local.abs.errormargin) +
  xlab(expression(lambda_c)) +
  facet_wrap(lbin ~ .)
```

## Mis-centering effect on scaling relations

```{r}
local.frac.errormargin =  frac.errormargin
local.abs.errormargin = abs.errormargin
```

```{r include = FALSE}
lo_lc <- read_tsv("../release-build/data/lo_lc_test.out")
lo_lc$Rbin <- as.factor(lo_lc$Rmis)
lo_lc <- mutate(lo_lc,
                 resid = probtrue - probtest,
                 fracerr = resid/probtrue,
                 lo = pmin(-local.frac.errormargin * probtrue, -local.abs.errormargin),
                 hi = pmax(local.frac.errormargin * probtrue, local.abs.errormargin),
                 within.frac = (abs(fracerr) < local.frac.errormargin),
                 within.abs = (abs(resid) < local.abs.errormargin),
                 pass = within.frac | within.abs)
```

Mis-centering cause a shift and an extra scatter to the values of cluster richness $\lambda$. $P(\lambda_o/\lambda_c | R_{mis})$ describes the probabilistic distribution of the richness shift ($\lambda_o/\lambda_c$) with a mis-centering distance $R_{mis}$ 
Our calculation is implemented in the file `lo_lc_t.hh`.

For this testing, we compare results from our implementation to the  independent function evaluations from Yuanyuan Zhang's mis-centering analysis (Zhang et al. 2019).

a. The comparisons are made at `r nunique(lo_lc$deltal)` unique values of $\lambda_{o}/\lambda_{c}$ with `r nlevels(lo_lc$Rbin)` different $R_{mis}$ values.
a. Over the range tested 
($`r min(lo_lc$deltal)` < \lambda_{cen} < `r max(lo_lc$deltal)`$),
the maximum magnitude of the absolute residual between the expected value
and our calculated value is
$`r max(abs(lo_lc$resid))`$;
the maximum magnitude of the fractional residual is
$`r max(abs(lo_lc$fracerr))`$.
a. For this test,
we require the relative differences to be 
less than $`r local.frac.errormargin`$,
or the absolute difference to be less than
$`r local.abs.errormargin`$.
a. `r testresult(lo_lc)`

Each panel in the figure below corresponds to a $R_{mis}$ value, specified in the panel top label.

```{r fig.asp=0.7, fig.cap = "Absolute comparison of our implementation to the expected results. Note that over-plotting may obscure the expected values, if the approximated values are very close."}
ggplot(lo_lc) +
  geom_smallpoint(aes(x = deltal , y = probtrue), color = truecolor, shape = trueshape) +
  geom_smallpoint(aes(x = deltal , y = probtest), color = testcolor, shape = testshape) +
  scale_y_log10() +
  scale_x_log10() +
  xlab(expression(lambda_miscent/lambda_cent)) +
  ylab("probability") +
  facet_wrap(Rbin~.)
```

```{r fig.asp=0.7, fig.cap = "Relative residual comparison of our implementation to the expected results. Blue points pass, and red points fail, the absolute residual test.", warning = FALSE}
ggplot(lo_lc, aes(x = deltal, y = fracerr, color = within.abs)) +
  geom_smallpoint() +
  guides(color = "none") +
  xlab(expression(lambda_miscent/lambda_cent)) +
  ylab("fractional residual") +
  geom_hline(yintercept = local.frac.errormargin, color = "blue") +
  geom_hline(yintercept = -1 * local.frac.errormargin, color = "blue") +
  ylim(-2 * local.frac.errormargin, 2 * local.frac.errormargin) +
  facet_wrap(Rbin~ .)
```

```{r fig.asp = 0.7, fig.cap = "Absolute residual comparison of our implementation to the expected results. Blue points pass, and red points fail, the relative residual test.", warning = FALSE}
ggplot(lo_lc, aes(x= deltal, y = resid, color = within.frac)) +
  geom_smallpoint() +
  guides(color = "none") +
  geom_hline(yintercept = local.abs.errormargin, color = "blue") +
  geom_hline(yintercept = -1 * local.abs.errormargin, color = "blue") +
  ylim(-2 * local.abs.errormargin, 2 * local.abs.errormargin) +
  xlab(expression(lambda_miscent/lambda_cent)) +
  facet_wrap(Rbin ~ .)
```

# Number Counts and Masses in SDSS

**This test requires running sdss_analysis.ini in cosmosis_tests first**.

```{r}
local.frac.errormargin =  0.05
local.abs.errormargin = abs.errormargin
```

In this section, we compare the stacked number counts and mass predictions in stacked richness/mass bins for a SDSS like analysis. These values are the integrated results using components listed above. For this testing, we compare results from our implementation to the implementations from Matteo Costanzi for the SDSS cosmology analysis (Costanzi et al. 2019). The test is conducted for the following set of cosmology and scaling relation parameters specified in **cosmosis_tests/values_comp.ini**.

```{r include = FALSE}
inte <- read_tsv("../data/integrated_bin_test.out")
inte <- mutate(inte,
               lambda = (llo+lhi)/2.0,
               residn = yntrue - yntest,
                 fracerrn = residn/yntrue,
                 residm = ymtrue - ymtest,
                 fracerrm = residm/ymtrue,
                 lo = pmin(-local.frac.errormargin * yntrue, -local.abs.errormargin),
                 hi = pmax(local.frac.errormargin * yntrue, local.abs.errormargin),
                 within.frac = (abs(fracerrn) < local.frac.errormargin),
                 within.abs = (abs(residn) < local.abs.errormargin),
                 pass = within.frac | within.abs)
```

a. The comparisons are made at `r nunique(inte$lambda)` unique values of $\lambda$ bins.
a. Over the **number of bins** tested and for the stacked **number counts**,
the maximum magnitude of the absolute residual between the expected value
and our calculated value is
$`r max(abs(inte$residn))`$;
the maximum magnitude of the fractional residual is
$`r max(abs(inte$fracerrn))`$.
a. Over the **number of bins** tested and for the stacked **average masses**,
the maximum magnitude of the absolute residual between the expected value
and our calculated value is
$`r max(abs(inte$residm))`$;
the maximum magnitude of the fractional residual is
$`r max(abs(inte$fracerrm))`$. 

```{r fig.asp=0.25, fig.cap = "Absolute comparison of our implementation to the expected results. Note that over-plotting may obscure the expected values, if the approximated values are very close."}
ggplot(inte) +
  geom_point(aes(x =lambda , y = yntrue), color = truecolor, shape = trueshape) +
  geom_point(aes(x = lambda , y = yntest), color = testcolor, shape = testshape) +
  scale_x_log10() +
  scale_y_log10() +
  xlab('Richness') +
  ylab("Number Counts")
```

```{r fig.cap = "Relative residual implementation or our approximation to the expected results. Points in blue pass, and points in red fail, the absolute residual test."}
ggplot(inte, aes(x = lambda, y = fracerrn, color = within.abs)) +
  geom_point() +
  guides(color = "none") +
  scale_x_log10() +
  xlab("richness") +
  ylab("fractional residual for number counts") +
  ylim(-5 * local.frac.errormargin, 5 * local.frac.errormargin) +
  geom_hline(yintercept = local.frac.errormargin, color="blue") +
  geom_hline(yintercept = -1 * local.frac.errormargin, color="blue")
```

```{r fig.asp=0.25, fig.cap = "Absolute comparison of our implementation to the expected results for average masses. Note that over-plotting may obscure the expected values, if the approximated values are very close."}
ggplot(inte) +
  geom_point(aes(x =lambda , y = ymtrue), color = truecolor, shape = trueshape) +
  geom_point(aes(x = lambda , y = ymtest), color = testcolor, shape = testshape) +
  scale_x_log10() +
  xlab('Richness') +
  ylab("LogMass")
```

```{r fig.cap = "Relative residual implementation or our approximation to the expected results for average masses. Points in blue pass, and points in red fail, the absolute residual test."}
ggplot(inte, aes(x = lambda, y = fracerrm, color = within.abs)) +
  geom_point() +
  scale_x_log10() +
  guides(color = "none") +
  xlab("richness") +
  ylab("fractional residual for average mass") +
   ylim(-2 * local.frac.errormargin, 2 * local.frac.errormargin) +
  geom_hline(yintercept = local.frac.errormargin, color="blue") +
  geom_hline(yintercept = -1 * local.frac.errormargin, color="blue")
```

# Number Counts and Masses in DESY1

**This test requires running y1_analysis.ini in cosmosis_tests first**

```{r}
local.frac.errormargin =  0.05
local.abs.errormargin = abs.errormargin
```

In this section, we compare the stacked number counts and mass predictions in stacked richness/mass bins for a y1 like analysis. These values are the integrated results using components listed above. For this testing, we compare results from our implementation to the implementations from Matteo Costanzi for the y1 cosmology analysis (DES 2019). The test is conducted for the following set of cosmology and scaling relation parameters specified in **cosmosis_tests/values_comp_y1.in**

```{r include = FALSE}
inte <- read_tsv("../data/integrated_bin_y1_test.out")
inte$zbin <- as.factor(inte$zlo)
inte <- mutate(inte,
               lambda = (llo+lhi)/2.0,
               residn = yntrue - yntest,
                 fracerrn = residn/yntrue,
                 residm = ymtrue - ymtest,
                 fracerrm = residm/ymtrue,
                 lo = pmin(-local.frac.errormargin * yntrue, -local.abs.errormargin),
                 hi = pmax(local.frac.errormargin * yntrue, local.abs.errormargin),
                 within.frac = (abs(fracerrn) < local.frac.errormargin),
                 within.abs = (abs(residn) < local.abs.errormargin),
                 pass = within.frac | within.abs)
```

a. The comparisons are made at `r nunique(inte$lambda)` unique values of $\lambda$ bins and `r nunique(inte$zlo)` unique values of $z$ bins.
a. Over the **number of bins** tested and for the stacked **number counts**,
the maximum magnitude of the absolute residual between the expected value
and our calculated value is
$`r max(abs(inte$residn))`$;
the maximum magnitude of the fractional residual is
$`r max(abs(inte$fracerrn))`$.
a. Over the **number of bins** tested and for the stacked **average masses**,
the maximum magnitude of the absolute residual between the expected value
and our calculated value is
$`r max(abs(inte$residm))`$;
the maximum magnitude of the fractional residual is
$`r max(abs(inte$fracerrm))`$. 

```{r fig.asp=0.3, fig.cap = "Absolute comparison of our implementation to the expected results. Note that over-plotting may obscure the expected values, if the approximated values are very close."}
ggplot(inte) +
  geom_point(aes(x =lambda , y = yntrue), color = truecolor, shape = trueshape) +
  geom_point(aes(x = lambda , y = yntest), color = testcolor, shape = testshape) +
  scale_x_log10() +
  xlab('Richness') +
  ylab("Number Counts")+
  facet_wrap(zbin~.)
```

```{r fig.asp=0.3, fig.cap = "Relative residual implementation or our approximation to the expected results. Points in blue pass, and points in red fail, the absolute residual test."}
ggplot(inte, aes(x = lambda, y = fracerrn, color = within.abs)) +
  geom_point() +
  scale_x_log10() +
  guides(color = "none") +
  xlab("richness") +
  ylab("fractional residual for number counts") +
   ylim(-2 * local.frac.errormargin, 2 * local.frac.errormargin) +
  geom_hline(yintercept = local.frac.errormargin, color="blue") +
  geom_hline(yintercept = -1 * local.frac.errormargin, color="blue")+
  facet_wrap(zbin~.)
```

```{r fig.asp=0.3, fig.cap = "Absolute comparison of our implementation to the expected results for average masses. Note that over-plotting may obscure the expected values, if the approximated values are very close."}
ggplot(inte) +
  geom_point(aes(x =lambda , y = ymtrue), color = truecolor, shape = trueshape) +
  geom_point(aes(x = lambda , y = ymtest), color = testcolor, shape = testshape) +
  scale_x_log10() +
  xlab('Richness') +
  ylab("LogMass")+
  facet_wrap(zbin~.)
```

```{r fig.asp=0.3, fig.cap = "Relative residual implementation or our approximation to the expected results for average masses. Points in blue pass, and points in red fail, the absolute residual test."}
ggplot(inte, aes(x = lambda, y = fracerrm, color = within.abs)) +
  geom_point() +
  scale_x_log10() +
  guides(color = "none") +
  xlab("richness") +
  ylab("fractional residual for average mass") +
   ylim(-2 * local.frac.errormargin, 2 * local.frac.errormargin) +
  geom_hline(yintercept = local.frac.errormargin, color="blue") +
  geom_hline(yintercept = -1 * local.frac.errormargin, color="blue")+
  facet_wrap(zbin~.)
```

# Cluster Mass Profile

The Cluster Mass profiles are calculated by a **separate CosmoSIS module**, which is a python wrapper to Tom McClintock's Cluster_toolkit code. The code exists in the folder `delta_sigma`.

**Before your run this test, make sure to go to the folder `deltasigma` and run
`cosmosis deltasigma_test.ini` to produce the proper test files.**
If you are using the supplied `GNUmakefile` to build the document,
this is done automatically.

## $\xi$ NFW

```{r}
local.frac.errormargin = 0.001
local.abs.errormargin = abs.errormargin
```

```{r include = FALSE}
xi_nfw <- read_tsv("../deltasigma/test/xi_nfwonly.out")
print(xi_nfw)
xi_nfw <- mutate(xi_nfw,
                 resid = ytrue - ytest,
                 fracerr = resid/ytrue,
                 lo = pmin(-local.frac.errormargin * ytrue, -local.abs.errormargin),
                 hi = pmax(local.frac.errormargin * ytrue, local.abs.errormargin),
                 within.frac = (abs(fracerr) < local.frac.errormargin),
                 within.abs = (abs(resid) < local.abs.errormargin),
                 pass = within.frac | within.abs)
```

$\xi_{NFW}(r)$ is the cluster matter correlation function in the 1-halo regime, adopting an NFW model. 

For this testing, we compare results from our implementation to the function evaluations from directly calling Tom McClintock's python/c package, for a cluster of $M_{200m}=3.199267137797384375e+14 M_\odot/h$ and $c=5$.

a.  The comparisons are made at `r nrow(xi_nfw)` unique values of $r$. 
a. Over the range tested 
($`r min(xi_nfw$r)` < r < `r max(xi_nfw$r)`$),
the maximum magnitude of the absolute residual between the expected value
and our calculated value is
$`r max(abs(xi_nfw$resid))`$;
the maximum magnitude of the fractional residual is
$`r max(abs(xi_nfw$fracerr))`$.
a. For this test,
we require the relative differences to be 
less than $`r local.frac.errormargin`$,
or the absolute difference to be less than
$`r local.abs.errormargin`$.
a. `r testresult(xi_nfw)`

The values of $\xi$ cover a very wide range, so a logarithmic $y$-axis is needed.
However, many points in the plot have $\xi = 0$.
In order to allow the use of a logarithmic axis, these points are suppressed.

```{r fig.cap = "Absolute comparison of our implementation to the expected results. Note that over-plotting may obscure the expected values, if the approximated values are very close. Note that we suppress negative values of $\\xi$ in this plot, so that we can use a log scale."}
ggplot(filter(xi_nfw, ytrue > 0.0, ytest > 0.0)) +
  geom_smallpoint(aes(x = r, y = ytrue), color = truecolor, shape = trueshape) +
  geom_smallpoint(aes(x = r, y = ytest), color = testcolor, shape = testshape) +
  scale_x_log10() +
  scale_y_log10() +
  xlab(expression(r)) +
  ylab(expression(xi))
```

```{r fig.cap = "Relative residual implementation or our approximation to the expected results. Points in blue pass, and points in red fail, the absolute residual test."}
ggplot(filter(xi_nfw, ytrue > 0.0, ytest > 0.0),
      aes(x = r, y = fracerr, color = within.abs)) +
  geom_smallpoint() +
  scale_x_log10() +
  guides(color = "none") +
  xlab(expression(r)) +
  ylab("fractional residual") +
  ylim(-2 * local.frac.errormargin, 2 * local.frac.errormargin) +
  geom_hline(yintercept = local.frac.errormargin, color="blue") +
  geom_hline(yintercept = -1 * local.frac.errormargin, color="blue")
```

## $\xi$ 2-halo

```{r}
local.frac.errormargin = 0.001
local.abs.errormargin = abs.errormargin
```

```{r include = FALSE}
xi_2h <- read_tsv("../deltasigma/test/xi_mm_2halo.out")
# We limit the range of R in our plots. It would be better to do this
# by controlling the generation of the xi_mm_2halo.out file.
print(xi_2h)
xi_2h <- mutate(xi_2h,
                 resid = ytrue - ytest,
                 fracerr = resid/ytrue,
                 lo = pmin(-local.frac.errormargin * ytrue, -local.abs.errormargin),
                 hi = pmax(local.frac.errormargin * ytrue, local.abs.errormargin),
                 within.frac = (abs(fracerr) < local.frac.errormargin),
                 within.abs = (abs(resid) < local.abs.errormargin),
                 pass = within.frac | within.abs)
```

$\xi_{mm}(r)$ is the matter correlation function in the cluster 2-halo regime, adopting a non-linear matter power spectrum, and not accounting for cluster bias.

For this testing, we compare results from our implementation to the function evaluations from directly calling Tom McClintock's python/c package, for $z=0.201010$.

a. The comparisons are done at `r nrow(xi_2h)` unique values of $r$
a. Over the range tested 
($`r min(xi_2h$r)` < r < `r max(xi_2h$r)`$),
the maximum magnitude of the absolute residual between the expected value
and our calculated value is
$`r max(abs(xi_2h$resid))`$;
the maximum magnitude of the fractional residual is
$`r max(abs(xi_2h$fracerr))`$.
a. For this test,
we require the relative differences to be 
less than $`r local.frac.errormargin`$,
or the absolute difference to be less than
$`r local.abs.errormargin`$.
a. `r testresult(xi_2h)`

The values of $\xi$ cover a very wide range, so a logarithmic $y$-axis is needed.
However, many points in the plot have $\xi = 0$.
In order to allow the use of a logarithmic axis, these points are suppressed.

```{r fig.cap = "Absolute comparison of our implementation to the expected results. Note that over-plotting may obscure the expected values, if the approximated values are very close. Note that we suppress negative values of $\\xi$ in this plot, so that we can use a log scale."}
ggplot(filter(xi_2h, ytest > 0.0, ytrue > 0.0)) +
  geom_smallpoint(aes(x = r, y = ytrue), color = truecolor, shape = trueshape) +
  geom_smallpoint(aes(x = r, y = ytest), color = testcolor, shape = testshape) +
  scale_x_log10() +
  scale_y_log10() +
  xlab(expression(r)) +
  ylab(expression(xi))
```

```{r fig.cap = "Relative residual comparison or our implementation to the expected results. Points in blue pass, and points in red fail, the absolute residual test. Note that points for which both the predicted and expected values are zero are suppressed."}
ggplot(filter(xi_2h, ytest > 0.0, ytrue > 0.0), 
       aes(x = r, y = fracerr, color = within.abs)) +
  geom_smallpoint() +
  scale_x_log10() +
  guides(color = "none") +
  #ylim(-2 * local.frac.errormargin, 2 * local.frac.errormargin) +
  xlab(expression(r)) +
  ylab("fractional residual") +
  geom_hline(yintercept = local.frac.errormargin, color="blue") +
  geom_hline(yintercept = -1 * local.frac.errormargin, color="blue")
```

## $\Sigma$ NFW

```{r}
local.frac.errormargin = 0.001
local.abs.errormargin = abs.errormargin
```

```{r include = FALSE}
sig_nfw <- read_tsv("../deltasigma/test/Sigma_nfwonly.out")
print(sig_nfw)
sig_nfw <- mutate(sig_nfw,
                 resid = ytrue - ytest,
                 fracerr = resid/ytrue,
                 lo = pmin(-local.frac.errormargin * ytrue, -local.abs.errormargin),
                 hi = pmax(local.frac.errormargin * ytrue, local.abs.errormargin),
                 within.frac = (abs(fracerr) < local.frac.errormargin),
                 within.abs = (abs(resid) < local.abs.errormargin),
                 pass = within.frac | within.abs)
```

$\Sigma_{NFW}(R)$ is the cluster surface mass density profile in the 1-halo regime, adopting an NFW model.

For this testing, we compare results from our implementation to the function evaluations from directly calling Tom McClintock's python/c package, for a cluster of $M_{200m}=3.199267137797384375e+14 M_\odot/h$ and $c=5$.

a. The comparisons are done at `r nrow(sig_nfw)` unique values of $R$.
a. Over the range tested 
($`r min(sig_nfw$r)` < R < `r max(sig_nfw$r)`$),
the maximum magnitude of the absolute residual between the expected value
and our calculated value is
$`r max(abs(sig_nfw$resid))`$;
the maximum magnitude of the fractional residual is
$`r max(abs(sig_nfw$fracerr))`$.
a. For this test,
we require the relative differences to be 
less than $`r local.frac.errormargin`$,
or the absolute difference to be less than
$`r local.abs.errormargin`$.
a. `r testresult(sig_nfw)`

```{r fig.cap = "Absolute comparison of our implementation to the expected results. Note that over-plotting may obscure the expected values, if the approximated values are very close."}
ggplot(sig_nfw) +
  geom_smallpoint(aes(x = r, y = ytrue), color = truecolor, shape = trueshape) +
  geom_smallpoint(aes(x = r, y = ytest), color = testcolor, shape = testshape) +
  scale_x_log10() +
  scale_y_log10() +
  xlab("R [Mpc/h]") +
  ylab(expression(Sigma))
```

```{r fig.cap = "Relative residual comparison or our approximation to the expected results. Points in blue pass, and points in red fail, the absolute residual test."}
ggplot(sig_nfw, aes(x = r, y = fracerr, color = within.abs)) +
  geom_smallpoint() +
  scale_x_log10() +
  guides(color = "none") +
  ylim(-2 * local.frac.errormargin, 2 * local.frac.errormargin) +
  xlab("R [Mpc/h]") +
  ylab("fractional residual") +
  geom_hline(yintercept = local.frac.errormargin, color="blue") +
  geom_hline(yintercept = -1 * local.frac.errormargin, color="blue")
```

## $\Sigma$ 2-halo

```{r}
local.frac.errormargin = 0.01
local.abs.errormargin = abs.errormargin
```

```{r include = FALSE}
sig_nfw <- read_tsv("../deltasigma/test/Sigma_mm_2halo.out")
print(sig_nfw)
# We cut off the range of R values at 30 Mpc. It would be better to do
# this by controlling the generation of the Sigma_mm_2halo.out file.
sig_nfw <- filter(sig_nfw, r <= 30)
sig_nfw <- mutate(sig_nfw,
                 resid = ytrue - ytest,
                 fracerr = resid/ytrue,
                 lo = pmin(-local.frac.errormargin * ytrue, -local.abs.errormargin),
                 hi = pmax(local.frac.errormargin * ytrue, local.abs.errormargin),
                 within.frac = (abs(fracerr) < local.frac.errormargin),
                 within.abs = (abs(resid) < local.abs.errormargin),
                 pass = within.frac | within.abs)
```

$\Sigma_{mm}(R)$ is the cluster surface mass density profile in the cluster 2-halo regime, adopting a non-linear matter power spectrum, and not accounting for cluster bias.

a. The comparisons are done at `r nrow(sig_nfw)` unique values of $R$.
a. Over the range tested 
($`r min(sig_nfw$r)` < R < `r max(sig_nfw$r)`$),
the maximum magnitude of the absolute residual between the expected value
and our calculated value is
$`r max(abs(sig_nfw$resid))`$;
the maximum magnitude of the fractional residual is
$`r max(abs(sig_nfw$fracerr))`$.
a. For this test,
we require the relative differences to be 
less than $`r local.frac.errormargin`$,
or the absolute difference to be less than
$`r local.abs.errormargin`$.
a. `r testresult(sig_nfw)`

```{r fig.cap = "Absolute comparison of our implementation to the expected results. Note that over-plotting may obscure the expected values, if the approximated values are very close."}
ggplot(sig_nfw) +
  geom_smallpoint(aes(x = r, y = ytrue), color = truecolor, shape = trueshape) +
  geom_smallpoint(aes(x = r, y = ytest), color = testcolor, shape = testshape) +
  scale_x_log10() +
  scale_y_log10() +
  xlab("R [Mpc/h]") +
  ylab(expression(Sigma))
```

```{r fig.cap = "Relative residual comparison or our approximation to the expected results. Points in blue pass, and points in red fail, the absolute residual test."}
ggplot(sig_nfw, aes(x = r, y = fracerr, color = within.abs)) +
  geom_smallpoint() +
  scale_x_log10() +
  guides(color = "none") +
  ylim(-2 * local.frac.errormargin, 2 * local.frac.errormargin) +
  xlab("R [Mpc/h]") +
  ylab("fractional residual") +
  geom_hline(yintercept = local.frac.errormargin, color="blue") +
  geom_hline(yintercept = -1 * local.frac.errormargin, color="blue")
```

## Halo bias $b(M_{200m}, z)$

```{r}
local.frac.errormargin =  0.001
local.abs.errormargin = abs.errormargin
```

```{r include = FALSE}
bias <- read_tsv("../deltasigma/test/bias.out")
bias$zbin <- as.factor(bias$z)
bias <- mutate(bias,
                 logmass  = log(M)/log(10),
                 resid = ytrue - ytest,
                 fracerr = resid/ytrue,
                 lo = pmin(-local.frac.errormargin * ytrue, -local.abs.errormargin),
                 hi = pmax(local.frac.errormargin * ytrue, local.abs.errormargin),
                 within.frac = (abs(fracerr) < local.frac.errormargin),
                 within.abs = (abs(resid) < local.abs.errormargin),
                 pass = within.frac | within.abs)
```

$b(M_{200m}, z)$ is the halo bias at mass $M_{200m}$ and redshift $z$.

For this testing, we compare results from our implementation to the function evaluations from directly calling Tom McClintock's python/c package.

a. The comparisons are done at `r nlevels(bias$zbin)` different $z$ values,
and `r nunique(bias$logmass)` different $M_{200m}$ values.
a. Over the range tested 
($`r min(bias$logmass)` < Log M_{200m} < `r max(bias$logmass)`$),
the maximum magnitude of the absolute residual between the expected value
and our calculated value is
$`r max(abs(bias$resid))`$;
the maximum magnitude of the fractional residual is
$`r max(abs(bias$fracerr))`$.
a. For this test,
we require the relative differences to be 
less than $`r local.frac.errormargin`$,
or the absolute difference to be less than
$`r local.abs.errormargin`$.
a. `r testresult(bias)`

Each panel in the figure below corresponds to a $z$ bin;
the value of $z$ is specified in the panel top label.

```{r fig.cap = "Absolute implementation of our approximation to the expected results. Note that over-plotting may obscure the expected values, if the approximated values are very close."}
ggplot(bias) +
  geom_point(aes(x = logmass , y = ytrue), color = truecolor, shape = trueshape) +
  geom_point(aes(x = logmass , y = ytest), color = testcolor, shape = testshape) +
  scale_y_log10() +
  xlab("Log M200m") +
  ylab("Bias") +
  facet_wrap(zbin~.)
```

```{r, fig.cap = "Relative residual implementation of our approximation to the expected results. Blue points pass, and red points fail, the absolute residual test.", warning = FALSE}
ggplot(bias, aes(x = logmass, y = fracerr, color = within.abs)) +
  geom_point() +
  guides(color = "none") +
  scale_y_log10() +
  xlab("Log M200m") +
  ylab("fractional residual") +
  geom_hline(yintercept = local.frac.errormargin, color = "blue") +
  geom_hline(yintercept = -1 * local.frac.errormargin, color = "blue") +
  ylim(-2 * local.frac.errormargin, 2 * local.frac.errormargin) +
  facet_wrap(zbin~ .)
```

# Mass Profile corrections

## Mis-centering offset

```{r}
local.frac.errormargin = frac.errormargin
local.abs.errormargin = abs.errormargin
```

```{r, include = FALSE}
miscent <- read_tsv("../release-build/data/Rmis_gammadist_test.out")
miscent <- mutate(miscent,
                  resid = probtrue - probtest,
                  fracerr = resid/probtrue,
                  lo = pmin(-local.frac.errormargin * probtrue, -local.abs.errormargin),
                  hi = pmax(local.frac.errormargin * probtrue, local.abs.errormargin),
                  within.frac = (abs(fracerr) < local.frac.errormargin),
                  within.abs = (abs(resid) < local.abs.errormargin),
                  pass = within.frac | within.abs)
```

$P(R_{mis})$ is the probabilistic distribution of the mis-centering offset $R_{mis}$.
Our calculation is implemented in the file `roffset_t.hh`.

For this testing, we compare our function output to the evaluations from Yuanyuan Zhang in the DES Y1 centering analysis (Zhang et al. 2019).

a. The comparisons are done at `r nrow(miscent)` unique $R_{mis}$ values.
a. Over the range tested
$(`r min(miscent$rmis)` < R_{mis} < `r max(miscent$rmis)`)$,
the maximum magnitude of the absolute residual between the expected value
and our approximation is
$`r max(abs(miscent$resid))`$;
the maximum magnitude of the fractional residual is
$`r max(abs(miscent$fracerr))`$.
a. For this test, we require the relative residuals to be less than
$`r local.frac.errormargin`$,
or the absolute residuals to be less than
$`r local.abs.errormargin`$.
a. `r testresult(miscent)`

```{r fig.cap = "Absolute comparison of our implementation to the expected results. Note that over-plotting may obscure the expected values, if the approximated values are very close."}
ggplot(miscent) +
  geom_point(aes(x = rmis, y = probtrue), color = truecolor, shape = trueshape) +
  geom_point(aes(x = rmis, y = probtest), color = testcolor, shape = testshape) +
  xlab(expression(R_mis)) +
  ylab(expression(Probability))
```

```{r fig.cap = "Relative residual comparison or our implementation to the expected results. Points in blue pass, and points in red fail, the absolute residual test."}
ggplot(miscent, aes(x = rmis, y = fracerr, color = within.abs)) +
  geom_point() +
  guides(color = "none") +
  ylim(-2 * local.frac.errormargin, 2 * local.frac.errormargin) +
  xlab(expression(R_mis)) +
  ylab("fractional residual") +
  geom_hline(yintercept = local.frac.errormargin, color="blue") +
  geom_hline(yintercept = -1 * local.frac.errormargin, color="blue")
```

# Lensing

## Critical density

```{r}
local.frac.errormargin =  0.0001
local.abs.errormargin = 1.0e-21
```

```{r include = FALSE}
sig_crit <- read_tsv("../release-build/data/Sigma_crit_test.out")
sig_crit$zbin <- as.factor(sig_crit$zl)
sig_crit <- mutate(sig_crit,
                 resid = ytrue - ytest,
                 fracerr = resid/(ytrue+as.numeric(ytrue==0)),
                 lo = pmin(-local.frac.errormargin * ytrue, -local.abs.errormargin, 0),
                 hi = pmax(local.frac.errormargin * ytrue, local.abs.errormargin),
                 within.frac = (abs(fracerr) < local.frac.errormargin),
                 within.abs = (abs(resid) < local.abs.errormargin),
                 pass = within.frac | within.abs)
```


$\Sigma_{crit}(z_l, z_s)$ is the inverse of the critical surface density of a lens system, which changes with the redshift of the lens $z_l$ and the redshift of the source galaxy $z_s$. Our calculation is implemented in the file `sigma_crit_inverse_t.hh`. 

For this testing, we compare our function output to the evaluations using the astropy package with a flat cosmology $\Omega_m=0.3$, $\Omega_b=0.06$, $h_0=0.7$.

a. The comparisons are done at, `r nunique(sig_crit$zs)` unique $z_s$ values, with `r nlevels(sig_crit$zbin)` different $z_l$ values
a. Over the range tested 
 ($`r min(sig_crit$zs)` < zs < `r max(sig_crit$zs)`$),
the maximum magnitude of the absolute residual between the expected value
and our calculated value is
$`r max(abs(sig_crit$resid))`$;
the maximum magnitude of the fractional residual is
$`r max(abs(sig_crit$fracerr))`$.
a. For this test,
we require the relative differences to be 
less than $`r local.frac.errormargin`$,
**or** the absolute difference to be less than
$`r local.abs.errormargin`$.
a. `r testresult(sig_crit)`

Each panel in the figure below corresponds to a $\lambda_{c}$ bin;
the value of the $\lambda_{c}$ mass is specified in the panel top label.

```{r fig.asp=0.7, fig.cap = "Absolute comparison of our implementation to the expected results. Note that over-plotting may obscure the expected values, if the approximated values are very close."}
ggplot(sig_crit) +
  geom_smallpoint(aes(x = zs, y = ytrue), color = truecolor, shape = trueshape) +
  geom_smallpoint(aes(x = zs , y = ytest), color = testcolor, shape = testshape) +
  xlab("z_s") +
  ylab("Density [Mpc^(-3)]") +
  facet_wrap(vars(zbin))
```

```{r fig.asp=0.7, fig.cap = "Absolute residual comparison of our implementation to the expected results. Blue points pass, and red points fail, the relative residual test.", warning = FALSE}
ggplot(sig_crit, aes(x= zs, y = resid, color = within.frac)) +
  geom_smallpoint() +
  guides(color = "none") +
  geom_hline(yintercept = local.abs.errormargin, color = "blue") +
  geom_hline(yintercept = -1 * local.abs.errormargin, color = "blue") +
  ylim(-2 * local.abs.errormargin, 2 * local.abs.errormargin) +
  xlab("z_s") +
  facet_wrap(vars(zbin))
```

```{r fig.asp=0.7, fig.cap = "Relative residual comparison or our implementation to the expected results. Points in blue pass, and points in red fail, the absolute residual test. In this plot, the relative residual is forced to be 0 when both the true and test values are 0, as is the case when zs <= zl"}
ggplot(sig_crit, aes(x = zs, y = fracerr, color = within.abs)) +
  geom_smallpoint() +
  guides(color = "none") +
  geom_hline(yintercept = local.frac.errormargin, color="blue") +
  geom_hline(yintercept = -1 * local.frac.errormargin, color="blue")+
  ylim(-2 * local.frac.errormargin, 2 * local.frac.errormargin) +
  xlab("z_s") +
  facet_wrap(vars(zbin))
```

# Session information

The following R package versions were used to prepare this document.

```{r}
sessionInfo()
```

