\documentclass[a4paper, 11pt]{article}
\usepackage[utf8]{inputenc}
\usepackage[left=18mm, right=18mm, top=22mm, bottom=22mm]{geometry}

\usepackage{amsmath,amsfonts,amssymb,bm}
\usepackage[english]{babel}
\usepackage{hyperref}
\usepackage{xcolor}
\usepackage{fancyhdr}
\usepackage{graphicx}
\usepackage{booktabs}
\usepackage{float}
\usepackage{multirow}

% Preferred style -------------------------------------------------------------
\definecolor{harvardcrimson}{HTML}{8C1D40}
\definecolor{softgrey}{HTML}{F3F3F3}
\definecolor{kwcol}{HTML}{1F4E79}
\definecolor{datablockcol}{HTML}{4B644A}

\hypersetup{
  colorlinks=true,
  linkcolor=harvardcrimson,
  citecolor=harvardcrimson,
  urlcolor=harvardcrimson,
  hypertexnames=false,
}

\pagestyle{fancy}
\fancyhf{}
\rhead{\textcolor{harvardcrimson}{\textbf{DES Y3 cluster pipeline modules}}}
\lhead{\textcolor{harvardcrimson}{\textbf{J.\,H.\,Esteves}}}
\cfoot{\thepage}
\renewcommand{\headrulewidth}{0.4pt}
\renewcommand{\footrulewidth}{0.4pt}
\setlength{\headheight}{14pt}
\renewcommand{\baselinestretch}{1.12}

% algorithm2e ----------------------------------------------------------------
\usepackage[ruled,vlined,linesnumbered]{algorithm2e}
\SetKwInput{KwData}{Datablock in}
\SetKwInput{KwResult}{Datablock out}
\SetKwInput{KwConfig}{Ini knobs}
\SetKwFor{PerZ}{for each}{do}{end}
\SetKwProg{Fn}{Function}{:}{}
\DontPrintSemicolon

% Custom commands ------------------------------------------------------------
\newcommand{\lob}{\lambda^{\rm ob}}
\newcommand{\ltr}{\lambda^{\rm tr}}
\newcommand{\zob}{z^{\rm ob}}
\newcommand{\dprj}{\Delta^{\rm prj}}
\newcommand{\xinl}{\xi_{\rm NL}}
\newcommand{\hMpc}{h^{-1}\,\mathrm{Mpc}}
\newcommand{\hMsun}{h^{-1}\,M_{\odot}}
\newcommand{\Sprj}{\Sigma^{\rm prj}}
\newcommand{\DSprj}{\Delta\Sigma^{\rm prj}}
\newcommand{\Bsmall}{B_{\rm small}}
\newcommand{\Blarge}{B_{\rm large}}
\newcommand{\modname}[1]{\textcolor{kwcol}{\texttt{#1}}}
\newcommand{\blk}[1]{\textcolor{datablockcol}{\texttt{#1}}}
\newcommand{\pyfile}[1]{{\small\texttt{#1}}}

\title{\vspace{-4ex}\huge \textbf{DES Y3 cluster-cosmology pipeline:}\\[0.5ex]
       \Large \textbf{module reference, numerical recipes, and
                      time\,/\,precision optimisations}}
\author{J.\,H.\,Esteves}
\date{2026-05-09}

\begin{document}
\maketitle

\begin{abstract}
This document is the technical reference for the
\modname{mock\_mcmc\_cp\_camb.ini} production pipeline: the union of the
Python and C++ modules that take cosmology + HOD parameters and
produce number counts, the one-halo tangential shear, the Costanzi-2026
two-halo projection bias, and a Gaussian likelihood against a
pre-baked mock datavector for MCMC sampling.  It documents the
numerical recipe of every module, the datablock contract each one
relies on, the relationship between the two shear deliverables
($g_t^{1h}$ and $g_t^{\rm prj}$), and the time\,/\,precision
trade-offs that reduced the per-sample cost from
\textbf{$1.71\,\mathrm{s}$ to $1.15\,\mathrm{s}$} during the April--May
2026 audit.  All modules live at
\texttt{/pscratch/sd/j/jesteves/y3\_cluster\_cpp} and are known to
produce C++\,$\leftrightarrow$\,Python agreement to $\lesssim 1.4\%$ on
$\Sprj, \DSprj$ across the full $R \in [0.1, 30]\,\hMpc$ wall grid.
Algorithmic workflows are written in \texttt{algorithm2e} pseudocode
throughout.
\end{abstract}

\tableofcontents
\vspace{1em}

% ============================================================================
\section{Pipeline topology}
\label{sec:topology}

The module sequence in \modname{cosmosis-models/mock\_mcmc\_cp\_camb.ini} is:

\begin{center}
\fbox{\parbox{0.95\linewidth}{\centering
\modname{consistency} $\to$ \modname{growth} $\to$ \modname{cp\_camb} $\to$
\modname{mf\_tinker} $\to$ \modname{halo\_model} $\to$
\modname{average\_sigma\_crit\_inv} $\to$ \modname{sel\_function} $\to$
\modname{NumCountsSel} $\to$ \modname{Shear1hSel} $\to$
\modname{b\_sel\_marg} $\to$ \modname{bsel} $\to$
\modname{red\_shear\_prj} $\to$ \modname{likelihoods}}}
\end{center}

\modname{mock\_mcmc\_cp\_camb.ini} is the production driver that
extends the \modname{smoke\_full\_sel\_cp\_camb.ini} module list with
the Gaussian \modname{likelihoods} module
(\pyfile{y3\_buzzard/likelihood\_cp.py}) reading the pre-baked
\pyfile{data/mock/mock\_dv\_cp\_camb.npz}.  Smoke runs set
\texttt{sampler = test} to probe a single sample; MCMC runs set
\texttt{sampler = emcee} (10 walkers, 50 steps, 10000 samples) and
vary a 3-parameter HOD subspace (\texttt{log10\_Mmin},
\texttt{log10\_M1}, \texttt{sigma\_lambda}) to verify end-to-end
closure -- posteriors must recover the fiducial HOD within their
$68\%$ intervals.

Each module reads a set of named sections from the CosmoSIS datablock
and writes its own section; the pipeline is pure-functional and there
are no hidden side channels.  Table~\ref{tab:module-roles} summarises
the role of each.

\begin{table}[H]
\centering
\small
\begin{tabular}{@{}l l l p{0.43\linewidth}@{}}
\toprule
Module & Type & Timing & Role \\
\midrule
\modname{consistency}            & Py (CSL)   & $3$\,ms  & normalise cosmology \\
\modname{growth}                 & Fortran (CSL) & $2$\,ms & $D(z)$, $f(z)$ \\
\modname{cp\_camb}               & Py (local)   & $8$\,ms & emulator $P(k,z)$ \\
\modname{mf\_tinker}             & Fortran (CSL) & $190$\,ms & Tinker 2008 HMF \\
\modname{halo\_model}            & Py (local)   & $155$\,ms & $b(M,z)$, $\xinl(r,z)$, optional lensing \\
\modname{average\_sigma\_crit\_inv} & Py (local) & $5$\,ms & $\langle\Sigma_{\rm crit}^{-1}\rangle(z_l)$ \\
\modname{sel\_function}          & Py (local)   & $225$\,ms & $S_{ij}(\ln M, z)$ tabulation \\
\modname{NumCountsSel}           & C++ (CUBA)   & $117$\,ms & $N_i$ cluster counts \\
\modname{Shear1hSel}      & C++ (CUBA)   & $90$\,ms & $N_i[g_t^{1h}](R)$ (see \S\ref{sec:shear_deliverables}) \\
\modname{b\_sel\_marg}           & C++          & $76$\,ms & $(P_1, I_1, J)$ operators \\
\modname{bsel}                   & Py (local)   & $18$\,ms & $\Bsmall, \Blarge$ close-form \\
\modname{red\_shear\_prj}        & C++          & $247$\,ms & projection $\Sprj, \DSprj, g_t^{\rm prj}$ \\
\modname{likelihoods}            & Py (local)   & $<1$\,ms & Gaussian $(d-m)^{\!\top}\!C^{-1}\!(d-m)$ \\
\midrule
\multicolumn{2}{r}{\textbf{total per sample:}} & \multicolumn{2}{l}{\textbf{$\sim 1.15$\,s} (login node, single thread)} \\
\bottomrule
\end{tabular}
\caption{Module inventory with roles and wall-time contributions at
the post-audit defaults (see \S\ref{sec:timing}). ``CSL'' means the
CosmoSIS standard-library module; ``local'' is shipped with this
repo.}
\label{tab:module-roles}
\end{table}

% ============================================================================
\section{Shared conventions}
\label{sec:conventions}

\paragraph{Units.}  All lengths are in $\hMpc$ (comoving, per
little-$h$).  $\Sigma, \Delta\Sigma$ are in $\hMsun\,\mathrm{pc}^{-2}$.
Halo bias is dimensionless.  Cosmology parameters follow CAMB
conventions: $\omega_m = \Omega_m h^2$, $\omega_b = \Omega_b h^2$,
$\log(10^{10} A_s)$, $n_s$, $\sigma_8$.

\paragraph{Richness bin convention.}  DES Y3 richness bin edges are
$[20, 30, 45, 60, 200]$ with arithmetic centres $\{25, 37.5, 52.5,
130\}$.  The centre $130$ replaces a legacy value of $100$; all C++
evaluators now hard-code the correct centre or read a
\modname{lob\_centers} ini override.  Photo-$z$ bins are $[0.20, 0.35]
\cup [0.35, 0.50] \cup [0.50, 0.65]$ with midpoints
$\{0.275, 0.425, 0.575\}$.

\paragraph{Wall grid layout.}  Evaluators that produce per-bin
observables read flat 1-D axis arrays from the ini
(\texttt{zo\_low}, \texttt{zo\_high}, \texttt{lambda\_bin}, and
optionally \texttt{radii}) and assemble them point-wise.  The length
of every axis is $N_{\rm grid} = N_{\zob} \cdot N_{\lob} \cdot N_R$;
for the smoke setups $N_{\rm grid} = 12$ (scalar observables) or
$120$ (radial observables with $N_R = 10$).

\paragraph{Mass-axis mass-shift (\texttt{mf\_tinker}).}  CosmoSIS
\modname{mf\_tinker} stores \blk{mass\_function/m\_h} in units of
$\Omega_m \cdot \hMsun$ (Komatsu CRL convention), not plain $\hMsun$.
The C++ \texttt{HMF\_t} constructor rescales the spline axis internally
at construction time, so every downstream evaluator queries with
raw $\ln M_{\rm phys}$.  Python consumers that read
\blk{mass\_function/m\_h} directly (\modname{bsel.py},
\modname{compare\_b\_sel\_py\_vs\_cpp.py}) explicitly rescale
\texttt{m\_h *= (Om - Onu)} on read.

\paragraph{Growth factor normalisation.}  The CosmoSIS
\modname{structure/growth\_factor} module publishes an un-normalised
$D(z)$ that equals $a$ at early times (so $D(0) \neq 1$).  Any
downstream consumer that needs $\sigma(M, z)/\sigma(M, 0)$
must divide out $D(0)$ explicitly.  The bias branch in
\modname{halo\_model\_cosmosis.py} does this.

\paragraph{$\sigma_8$ effective.}  The \modname{cp\_camb} v2c emulator
caps $\log(10^{10} A_s)$ at $3.184$; with our fiducial
$\log(10^{10} A_s) = 3.97$ this clips and realises
$\sigma_8(z{=}0) \simeq 0.800$ regardless of the \texttt{sigma8\_input}
knob.  Python validations are run at the same effective
$\sigma_8 = 0.80$ to stay apples-to-apples; see the
\modname{sigma\_prj\_py\_vs\_cpp\_report.tex} companion for the audit.

% ============================================================================
\section{Observables and the shear composition}
\label{sec:shear_deliverables}

The pipeline emits two observables that the likelihood compares
against data:

\begin{enumerate}
  \setlength\itemsep{2pt}
  \item \blk{numcountssel/vals} -- 12 cluster-count bins, $N_i[1]$.
  \item Summed tangential shear, length 120 ($12 \times 10\,R$):
    \begin{equation}
    \label{eq:gt_sum}
      \gamma_t^{\rm theory}(R\,|\,i, j)
      \;=\; \langle\gamma_t^{1h}\rangle_i(R)
          \;+\; \gamma_t^{\rm prj}(R\,|\,\lambda_{\rm ob}, z_{\rm ob})
      \;=\; \frac{\blk{shear1hsel/vals}}{\blk{numcountssel/vals}}
          \;+\; \blk{shear\_prj/vals}.
    \end{equation}
\end{enumerate}

The pipeline also publishes \blk{sigma\_prj/vals},
\blk{dsigma\_prj/vals}, and the \texttt{rnd}/\texttt{cl} subfields of
\blk{shear\_prj/vals} for diagnostics.

\paragraph{Why the sum is linear.}  Both modules emit tangential shear
\begin{equation}
  \gamma_t(R) \;=\; \Delta\Sigma(R)\,\Sigma_{\rm crit}^{-1}(z),
\end{equation}
\emph{not} the reduced shear $g_t = \gamma_t/(1 - \Sigma \Sigma_{\rm crit}^{-1})$.
The earlier reduced-shear form coupled the two pieces through its
$1/(1-x)$ denominator, which makes $g_t^{\rm obs}$ nonlinear in the
$\Delta\Sigma$ decomposition and blocks the additive split.  Dropping
the denominator (retired 2026-05-11 in
\pyfile{src/modules/num\_counts\_sel/lensing\_weights.hh} and
\pyfile{src/models/sigma\_prj\_t.hh}) recovers the linear-in-$\Delta\Sigma$
form and makes Eq.~\eqref{eq:gt_sum} exact.

\subsection{The $N_i[f]$ operator}

\modname{NumCountsSel}, \modname{Shear1hSel},
\modname{MassWeightedSel}, \modname{BiasWeightedSel} all share the
same template \texttt{NOperatorSelRadial<F>} /
\texttt{NOperatorSelScalar<F>}
(\pyfile{src/models/n\_operator\_sel\_t.hh}).  Each instantiation
plugs a different integrand weight $F$ into the generic operator:
\begin{equation}
\label{eq:n_op}
  N_i[f](R)
  \;=\; \int d\ln M \int dz\
    \Omega(z)\,\frac{dV}{d\Omega\,dz}\,n(M, z)\,
    S_{ij}(\ln M, z)\,f(R, \ln M, z),
\end{equation}
with the richness-selection tensor $S_{ij}(\ln M, z) = S_i(\ln M, z)
K_j(z)$ published by \modname{sel\_function} (\S\ref{sec:sel_function}).
Choosing the integrand weight $f$ picks the observable:
\begin{center}
\small
\begin{tabular}{@{}l l p{0.45\linewidth}@{}}
\toprule
Weight $f$                     & Module                       & Observable built                                  \\
\midrule
$1$                            & \modname{NumCountsSel}       & $N_i$ cluster number count                        \\
$M$                            & \modname{MassWeightedSel}    & $\langle M\rangle_i = N_i[M]/N_i$                 \\
$b(M, z)$                      & \modname{BiasWeightedSel}    & $\langle b\rangle_i = N_i[b]/N_i$                 \\
$\gamma_t^{1h}(R; M, z)$       & \modname{Shear1hSel}         & $N_i[\gamma_t^{1h}](R)$ (see below)               \\
\bottomrule
\end{tabular}
\end{center}

\subsection{\texttt{Shear1hSel} is $N_i$-weighted; divide by \texttt{NumCountsSel}}
\label{sec:shear1h_vs_prj}

Key point: \modname{Shear1hSel} is \emph{not} the per-cluster 1-halo
shear $\gamma_t^{1h}(R)$.  It is the \emph{integral of the 1-halo
shear over the cluster population in the bin}, weighted by
$\Omega(z)\,dV/d\Omega/dz\,n(M, z)\,S_{ij}$.  Inside the template,
the weight is
\begin{equation}
  \gamma_t^{1h}(R; M, z) \;=\; \Delta\Sigma_{\rm NFW}(R, M)\,
                               \Sigma_{\rm crit}^{-1}(z),
\end{equation}
built from the \emph{centred} NFW profile
\blk{haloModel/dSigma\_nfw}.  To recover the per-cluster average one
divides by the number-count integral on the same wall point:
\begin{equation}
  \langle\gamma_t^{1h}\rangle_i(R)
  \;=\; \frac{N_i[\gamma_t^{1h}](R)}{N_i[1]}
  \;=\; \frac{\blk{shear1hsel/vals}}{\blk{numcountssel/vals}}.
\end{equation}
This division lives in the likelihood
(\pyfile{y3\_buzzard/likelihood\_cp.py}), not inside a single module.

\subsection{\texttt{shear\_prj} produces a different integrand}

\modname{shear\_prj} does \emph{not} pass through the $N_i[f]$
template.  Its integrand is
\begin{equation}
\label{eq:sigma_prj}
  \Sigma^{\rm prj}(R \,|\, \lambda_{\rm ob}, z_{\rm ob})
  \;=\; \int dz \int d\ln M\ \frac{dV}{d\Omega dz}\,w_z(z, z_{\rm ob})\,
    n(M, z)\,
    \big[1 + b(M, z)\,b_{\rm sel}(\theta)\,\xi_{\rm NL}\big]\,
    \Sigma_{\rm mis}(R \,|\, M, z, \theta D_A(z_{\rm ob})),
\end{equation}
with three structural differences vs Eq.~\eqref{eq:n_op}:

\begin{itemize}
  \setlength\itemsep{2pt}
  \item \textbf{No $S_{ij}$.}  A parabolic photo-$z$ kernel
    $w_z(z, z_{\rm ob}) = \max(0, 1 - u^2)$ replaces the richness
    selection tensor; the module evaluates \emph{per}
    $(\lambda_{\rm ob}, z_{\rm ob})$ wall point, already per
    richness bin.
  \item \textbf{No $\Omega(z)$.}  The survey-area factor cancels between
    $\Sigma^{\rm prj}$ and its normalisation, so it is not applied
    here.
  \item \textbf{Off-centred NFW $\Sigma_{\rm mis}$.}  The 1-halo
    density uses a \texttt{single}-offset miscentered NFW (offset by
    $R_{\rm mis} = \theta D_A(z_{\rm ob})$), \emph{not} the centred
    $\Sigma_{\rm NFW}$ used by \modname{Shear1hSel}.
\end{itemize}

Both the $[1]$ term (\texttt{rnd}) and the $[b\,b_{\rm sel}\,\xi_{\rm NL}]$
term (\texttt{cl}) are accumulated separately.  The same kernel
construction produces $\Delta\Sigma^{\rm prj}$ and the shear
$\gamma_t^{\rm prj} = \Delta\Sigma^{\rm prj} \Sigma_{\rm crit}^{-1}$
in one pass.

\subsection{Putting it together: $\gamma_t^{\rm theory} = \langle\gamma_t^{1h}\rangle + \gamma_t^{\rm prj}$}

The two modules integrate \emph{different kernels}
(\mbox{centred NFW} vs \mbox{miscentered $\Sigma_{\rm mis}$}) against
\emph{different weights} ($S_{ij}$ vs $w_z$).  In the language of the
Costanzi-2026 paper:
\begin{itemize}
  \setlength\itemsep{1pt}
  \item $\langle\gamma_t^{1h}\rangle_i(R)$ is the one-halo lensing
    signal expected for a perfectly-centred cluster population.
  \item $\gamma_t^{\rm prj}(R)$ is the \emph{projection-effect
    correction}: the additional shear contributed by miscentering
    plus correlated large-scale structure, integrated over the same
    cluster sample via the photo-$z$ kernel $w_z$.
\end{itemize}
The observable used to fit data is the sum of the two,
Eq.~\eqref{eq:gt_sum}.  The sum is \emph{exact} (not a leading-order
approximation) because both pieces are linear in $\Delta\Sigma$; the
old reduced-shear $1/(1-\Sigma\Sigma_{\rm crit}^{-1})$ coupling has
been dropped.  \modname{likelihood\_cp.py} assembles the summed
theory vector internally and compares it against the single
\texttt{data\_Shear} entry of the mock npz.

% ============================================================================
\section{\texttt{cp\_camb}: emulated linear $P(k, z)$}
\label{sec:cp_camb}

\modname{Path}\,\pyfile{src/modules/cp\_camb/cp\_camb.py}. A Python
wrapper around a CosmoPower neural-network emulator that predicts
$P_{\rm lin}(k, z{=}0)$ and reconstructs $P(k, z>0)$ via the growth
factor.  Publishes both total-matter and nonu (CDM+baryon) spectra.

\begin{algorithm}[H]
\caption{\texttt{cp\_camb.execute}}
\KwConfig{\texttt{linear\_pk\_path}, \texttt{linear\_nonu\_pk\_path},
\texttt{zmin}, \texttt{zmax}, \texttt{nz}, \texttt{apply\_growth}}
\KwData{\blk{cosmological\_parameters}, \blk{growth\_parameters/\{z,d\_z\}}}
\KwResult{\blk{matter\_power\_lin/\{z,k\_h,p\_k\}}, \blk{cdm\_baryon\_power\_lin/\{z,k\_h,p\_k\}}, \blk{distances/\{z,d\_a,d\_m,d\_l,h,mu\}}}

Read cosmology vector $\phi = \{\omega_m, \omega_b, h, n_s, \log(10^{10} A_s), m_\nu\}$\;
$D(z) \leftarrow$ \blk{growth\_parameters/d\_z}, \textbf{divide by $D(0)$}\;
Build linear CosmoPower forward pass: $\hat P_0(k) \leftarrow \text{emu}(\phi, z{=}0)$\;
\For{$z_i$ in linspace(\texttt{zmin}, \texttt{zmax}, \texttt{nz})}{
  $P(k, z_i) \leftarrow D(z_i)^2 \cdot \hat P_0(k)$\;
}
Write total-matter and nonu spectra to their respective sections\;
Write \texttt{distances/*} via \texttt{astropy.cosmology.FlatLambdaCDM}\;
\end{algorithm}

\paragraph{Key points.}
\begin{itemize}
  \setlength\itemsep{2pt}
  \item The emulator is z=0-only; z-evolution is reconstructed by
    $D(z)^2$ scaling.  Training-box clip on $\log(10^{10} A_s)$ is
    permanent --- $\sigma_8$ above $\sim 0.82$ is unreachable.
  \item The nonu spectrum is published under the \blk{cdm\_baryon\_power\_lin}
    section, which is what \modname{mf\_tinker} reads when
    \texttt{matter\_power\_lin\_version = 2} (the default in this
    pipeline).
  \item Distances are computed in astropy (not via the emulator),
    since \modname{cp\_camb} predicts only $P(k)$.  This keeps the two
    outputs from drifting with emulator training noise.
\end{itemize}

% ============================================================================
\section{\texttt{halo\_model}: bias $b(M,z)$ and $\xinl(r,z)$}
\label{sec:halo_model}

\modname{Path}\,\pyfile{y3\_buzzard/halo\_model\_cosmosis.py}. The
single most configurable module in the pipeline: it publishes the
\blk{haloModel/bias} and \blk{xi\_nl} tables that half the downstream
modules rely on, plus an \emph{optional} set of one-halo and two-halo
lensing profiles (\blk{Sigma\_nfw}, \blk{dSigma\_nfw}, \blk{Sigma\_hh},
\blk{dSigma\_hh}, \blk{Wp\_hh}).

\subsection{Algorithm}

\begin{algorithm}[H]
\caption{\texttt{halo\_model\_cosmosis.execute}}
\KwConfig{\texttt{R\_perp\_min}, \texttt{R\_perp\_max},
 \texttt{R\_perp\_bins}, \texttt{M\_min}, \texttt{M\_max}, \texttt{M\_bins},
 \texttt{compute\_lensing\_1h} (bool), \texttt{compute\_lensing\_2h} (bool)}
\KwData{\blk{matter\_power\_lin}, \blk{growth\_parameters}, \blk{cosmological\_parameters}}
\KwResult{\blk{haloModel/\{m\_h, lnM, z, bias, rhoc\}},
 \blk{xi\_nl/\{r,z,xi\_nl\}}, optionally
 \blk{haloModel/\{Sigma\_nfw, dSigma\_nfw, Sigma\_hh, dSigma\_hh, Wp\_hh, concentration, ...\}}}

Build mass grid $M \leftarrow$ logspace(\texttt{M\_min}, \texttt{M\_max}, \texttt{M\_bins})\;
Compute peak height $\nu(M) = \delta_c / \sigma(M, z{=}0)$ via \texttt{cluster\_toolkit.peak\_height.nu\_at\_M}\;
\For{each $z_i$ in \blk{matter\_power\_lin/z}}{
  $b(M, z_i) \leftarrow$ Tinker 2010 formula at $\nu(M) / (D(z_i)/D(0))$\;
  $\xinl(r, z_i) \leftarrow$ \texttt{ct.xi.xi\_mm\_at\_r} over $r \in [10^{-3}, 10^3]\,\hMpc$ (128 nodes)\;
}
Write \blk{haloModel/bias}, \blk{xi\_nl/xi\_nl}\;
\If{\texttt{compute\_lensing\_1h}}{
  NFW analytic $\Sigma(R, M)$, $\Delta\Sigma(R, M)$ via Child-18 concentration\;
  Write \blk{haloModel/\{Sigma\_nfw, dSigma\_nfw, concentration\}}\;
}
\If{\texttt{compute\_lensing\_2h}}{
  \texttt{ct\_2hTerm.pk\_to\_dsigma} (Hankel $P \to \xi \to \Sigma \to \Delta\Sigma$)\;
  Write \blk{haloModel/\{Sigma\_hh, dSigma\_hh, Wp\_hh\}}\;
}
\end{algorithm}

\subsection{Optimisation: split \texttt{compute\_lensing\_1h} / \texttt{compute\_lensing\_2h}}

The original module always ran both lensing branches.  The 2h branch
\texttt{ct\_2hTerm.pk\_to\_dsigma} is a redshift loop of
cluster\_toolkit Hankel transforms; at $N_z = 50$ it costs
$\sim 200$--$300\,$ms/sample and accounts for the majority of the
module's runtime.  \modname{red\_shear\_prj} (the modern 2-halo projection
module) does \emph{not} read the 2h outputs; only the legacy
\modname{ShearTotSel} family does.

\paragraph{Fix.}  Split the flag into \texttt{compute\_lensing\_1h}
(Shear1hSel needs it) and \texttt{compute\_lensing\_2h}
(nothing in \modname{smoke\_full\_sel\_cp\_camb.ini} reads it).  For this
pipeline:
\begin{verbatim}
  compute_lensing_1h = T
  compute_lensing_2h = F    # saves 180-200 ms
\end{verbatim}
Pure skip, no accuracy cost.

\paragraph{Growth-factor normalisation bug.}  Prior to May 2026 the
bias line was \texttt{bias\_at\_nu(nu / dazi)} where \texttt{dazi =
D(z)} un-normalised.  At $z{=}0.425$, $D(0) \simeq 0.758$, so the
effective $\nu$ was inflated by $1/D(0) \simeq 1.32$, producing halo
biases up to $2\times$ too large.  Fixed by dividing out
$D(0)$ before the scale.

% ============================================================================
\section{\texttt{sel\_function}: tabulate $S_{ij}(\ln M, z)$}
\label{sec:sel_function}

\modname{Path}\,\pyfile{src/modules/sel\_function/sel\_function.py}.
Pre-tabulates the richness-selection function on a shared
$(\ln M, z)$ grid that the C++ \modname{NumCountsSel} and
\modname{Shear1hSel} downstream modules serve via \texttt{Interp2D}.

\subsection{Mathematical form}

The published tensor is
\begin{equation}
  S_{ij}(\ln M, z) = S_i(\ln M, z) \cdot K_j(z)
                   = \Big[ \int d\ltr\, K_i(\ltr, z)\, P_{\rm HOD}(\ltr | M, z) \Big]
                     \cdot K_j(z),
\end{equation}
with the richness kernel $K_i$ computed as a closed-form CDF
difference of the Costanzi EMG $P(\lob | \ltr, z)$:
\begin{equation}
  K_i(\ltr, z) = F_{\rm EMG}(\lambda_{\rm max}^i \,|\, \ltr, z)
               - F_{\rm EMG}(\lambda_{\rm min}^i \,|\, \ltr, z),
\end{equation}
the photo-$z$ kernel $K_j$ a Gaussian CDF difference
\begin{equation}
  K_j(z) = \Phi\!\left( (z_{\rm max}^j - z)/\sigma_z \right)
         - \Phi\!\left( (z_{\rm min}^j - z)/\sigma_z \right),
\end{equation}
and $P_{\rm HOD}$ the continuous shifted-Poisson form
\begin{equation}
  P_{\rm HOD}(\ltr | M, z)
  = \exp\!\big[ -\nu + (\ltr + \delta - 1)\log\nu
               - \log\Gamma(\ltr + \delta) \big],
\end{equation}
with $\nu = \mu_{\rm sat}(M,z) + \delta$ and $\delta = (\sigma_{\rm intr} \mu_{\rm sat})^2$.

\subsection{Algorithm}

\begin{algorithm}[H]
\caption{\texttt{sel\_function.execute}}
\KwConfig{\texttt{lam\_min}, \texttt{lam\_max}, \texttt{zob\_min}, \texttt{zob\_max},
 \texttt{sigma\_z}, \texttt{n\_lnm}, \texttt{n\_z\_shared}, \texttt{N\_q}, \texttt{L\_lam}}
\KwData{\blk{cluster\_mor}, \blk{plob\_ltr\_params/*} or pre-baked npz fallback}
\KwResult{\blk{sel\_function/\{lnM, z, S\_stack\}}, shape $(N_{\rm bin}, N_z, N_{\ln M})$}

Build shared grids $\ln M \in [\texttt{lnm\_low}, \texttt{lnm\_high}]$ (linear, $N_{\ln M}$),
$z \in [\texttt{zt\_low}, \texttt{zt\_high}]$ (linear, $N_z$)\;
\tcp*[l]{Step 1: GL bracket + $P_{\rm HOD}$ on $(N_{\ln M}, N_z, N_q)$ tensor}
\For{\rm each $(\ln M_a, z_b)$}{
  Compute $\mu_{\rm sat}$, $\sigma_{\rm eff}$, and GL bracket
    $[\mu_{\rm eff} - L_\lambda \sigma_{\rm eff}, \mu_{\rm eff} + L_\lambda \sigma_{\rm eff}]$\;
  Place $N_q$ GL nodes $\ltr_k$ with weights $W_k$\;
}
Evaluate $P_{\rm HOD}(\ltr_k | \ln M_a, z_b)$ in one broadcast \texttt{gammaln} call\;
\tcp*[l]{Step 2: $K_i$ via CDF differencing at unique bin edges}
$\mu_p, \sigma_p, \tau_p, f_{{\rm prj},p} \leftarrow$ \texttt{plob\_params}$(\ltr_k, z)$
  \textbf{(1-D z fast path)}\;
$\{u_e\}$ = unique bin edges (e.g. $\{20, 30, 45, 60, 200\}$)\;
\For{\rm each edge $u_e$}{
  $F_e(\ltr_k, z) \leftarrow F_{\rm EMG}(u_e | \ltr_k, z)$ via \texttt{erfcx}\;
}
\tcp*[l]{Step 3: per-bin assembly}
\For{\rm each bin $j$}{
  $K_i \leftarrow F_{\rm edges[hi_j]} - F_{\rm edges[lo_j]}$\;
  $S_i(\ln M, z) = \sum_k W_k \cdot K_i(\ltr_k, z) \cdot P_{\rm HOD}(\ltr_k | \ln M, z)$\;
  $K_j(z) = \Phi((z_{\rm max}^j - z)/\sigma_z) - \Phi((z_{\rm min}^j - z)/\sigma_z)$\;
  \blk{sel\_function/S\_stack}$[j] = S_i \cdot K_j$\;
}
\end{algorithm}

\subsection{Optimisations}

\paragraph{1-D $z$ fast path in \texttt{\_plob\_params}.}  The 8 EMG
spline coefficients $a_\mu(z), b_\mu(z), \dots, b_{f_{\rm prj}}(z)$
depend only on $z$.  The original code evaluated each spline on a
pre-broadcast $(N_{\ln M}, N_z, N_q) = (256, 64, 32)$ tensor
(524\,k elements), at $\sim 130$\,ms total.  The fast path detects
when the ``$z$'' argument is 1-D of length $N_z$, evaluates the 8
splines on the 64-node grid, and reshapes the result to
$(1, N_z, 1)$ for numpy broadcasting.  \textbf{Saves $\sim 100$\,ms.}
Bit-identical numerical result.

\paragraph{CDF differencing at unique edges.}  Four richness bins
share five edges; computing \texttt{F\_EMG} at each edge once and
differencing costs 5 \texttt{erfcx} calls vs the naive 8 (four bins
$\times$ two endpoints each).

\paragraph{$n_{\ln M}$ grid sweep.}  A dedicated whole-pipeline
convergence study (\S\ref{sec:timing}, Tab.~\ref{tab:nlnm-sweep})
found that \texttt{n\_lnm = 192} minimises the \emph{total} pipeline
time, not just \modname{sel\_function} in isolation: reducing
$N_{\ln M}$ below 192 forces \modname{NumCountsSel}'s Cuhre adaptive
integrator to refine more, and the Cuhre overhead eats the
\modname{sel\_function} savings.  The ini is set to the whole-pipeline
optimum.

% ============================================================================
\section{\texttt{b\_sel\_marg}: $(P_1, I_1, J)$ operators}
\label{sec:b_sel_marg}

\modname{Path}\,\pyfile{src/models/p\_operator\_t.hh} +
\pyfile{src/modules/b\_sel\_marg\_cpu/BSelMargIntegrand.cc}. Fixed-GL
C++ evaluator that co-computes the three Costanzi-2026 $P[X]$
scalars on the $(\zob, \lob)$ wall.

\subsection{Mathematical form}

\begin{align}
  P_1(\lob, \zob) &= \int dz\, d\ln M\, d\ltr\, d\theta\,
                     \mathcal W \cdot f_A(\theta, \theta_\lambda(\ltr)), \\
  I_1(\lob, \zob) &= \int dz\, d\ln M\, d\ltr\, d\theta\,
                     \mathcal W \cdot b(M,z) \cdot \xinl(|\Delta \chi|, \zob)
                     \cdot \sigma(\theta) \cdot f_A, \\
  J(\lob, \zob)   &= I_2 - I_1 =
                     \int dz\, d\ln M\, d\ltr\, d\theta\,
                     \mathcal W \cdot b(M,z) \cdot \xinl \cdot (1-\sigma(\theta))
                     \cdot f_A.
\end{align}
The common integrand factor is
$\mathcal W = w_z(z) \cdot dV/d\Omega dz \cdot \mathbb 1[\theta > \theta_{\rm excl}(z)]
             \cdot dn/d\ln M \cdot P_{\rm HOD}(\ltr | M, z) \cdot \ltr
             \cdot 2\pi \sin\theta$.
$\sigma(\theta)$ is the Costanzi-2026 sigmoid
$\sigma = [1 + e^{-k(\theta - \theta_0)}]^{-1}$
with $k = 2.5/\theta_\lambda$, $\theta_0 = \theta_\lambda / 2$.

\subsection{Algorithm (theta-outer form)}

\begin{algorithm}[H]
\caption{\texttt{P\_operator::evaluate(lob, zob)}}
\KwConfig{\texttt{n\_theta}, \texttt{n\_zring}, \texttt{n\_zouter},
 \texttt{n\_lnm}, \texttt{n\_lt}, \texttt{R\_max\_cMpch}}
\KwData{\blk{distances}, \blk{mass\_function}, \blk{haloModel/bias}, \blk{xi\_nl}, \blk{cluster\_mor}}
\KwResult{\blk{b\_sel\_marg\_P1/vals}, \blk{b\_sel\_marg\_I1/vals}, \blk{b\_sel\_marg\_J/vals}}

\tcp*[l]{Build one $\theta$-grid per $(\lob, \zob)$}
$\theta_\lambda = R_\lambda(\lob) (1+\zob) / \chi(\zob)$\;
$\theta_{\rm lo} = \min_z \theta_{\rm excl}(z)$,\; $\theta_{\rm hi} = 2\theta_\lambda$\;
Split at $\theta_\lambda$: $\texttt{n\_theta}/2$ GL nodes per segment\;

\tcp*[l]{Pre-compute $\theta$-only caches}
\For{each $\theta_i$}{
  $\sin\theta_i, \cos\theta_i, \sigma(\theta_i), \omega_i = w_\theta \cdot 2\pi \sin\theta_i$\;
}

\tcp*[l]{Pre-contract mass: W\_P1, W\_I per (lt, iz)}
\For{$iz = 1, \dots, N_z$}{
  \If{$w_z(z_{iz}) = 0$}{continue\;}
  \For{$il = 1, \dots, N_{\rm lt}$}{
    $W_{P_1}[il, iz] = w_z \cdot dV \cdot \ltr[il] \cdot w_{\ltr}[il] \cdot \sum_M \texttt{HMF} \cdot p_{\rm MOR} \cdot w_M$\;
    $W_I[il, iz]     = W_{P_1}[il, iz]\ \text{weighted additionally by}\ b(M,z)$\;
  }
}

\tcp*[l]{Outer theta, inner z, innermost lt}
$\text{sum}_{P_1} = \text{sum}_{I_1} = \text{sum}_J = 0$\;
\For{$it = 1, \dots, N_\theta$}{
  \For{$iz = 1, \dots, N_z$}{
    \If{$\theta_{it} \le \theta_{\rm excl}(z_{iz})$}{continue\;}
    $\xi \leftarrow \xinl(|\Delta\chi(\theta_{it}, z_{iz})|, \zob)$\;
    \For{$il = 1, \dots, N_{\rm lt}$}{
      $f_A = \texttt{area\_overlap}(\theta_{it}, \theta_\lambda, \theta_\lambda^{\rm tr}(\ltr_{il}, z_{iz}))$\;
      acc\_$P_1$ += $f_A \cdot W_{P_1}[il, iz]$\;
      acc\_$I_1$ += $f_A \cdot W_I[il, iz] \cdot \xi \cdot \sigma(\theta_{it})$\;
      acc\_$J$   += $f_A \cdot W_I[il, iz] \cdot \xi \cdot (1-\sigma(\theta_{it}))$\;
    }
  }
  sum += $\omega_{it} \cdot $ acc\;
}
\end{algorithm}

\subsection{Key recipes}

\paragraph{Emit $J = I_2 - I_1$ directly, not the difference.}
As $\theta \to 2\theta_\lambda$, $\sigma(\theta) \to 1$, so the
$I_1$ and $I_2$ integrands become nearly equal: the GL sum for
$I_2 - I_1$ loses $\sim 1$ digit at large-$\theta$ nodes to
cancellation.  The integrand for $J$ is $(1-\sigma(\theta)) \xinl b\,
f_A$, always non-negative, no cancellation.  The downstream
$b_{\rm zero}$ formula \texttt{((lob-ltr) - P1 - b\_infty*I1) / J}
uses $J$ in the denominator, so emitting it directly also protects
the downstream arithmetic from subtractive noise.

\paragraph{$\theta$-outer restructure.}  Before the refactor the loop
nest was $z$-outer: $\theta$ and $\sigma(\theta)$ caches were rebuilt
$N_z \times$ per sample.  The new form hoists $\theta$ to the outer
axis (one \texttt{gl\_nodes} call per $(\lob, \zob)$, one
\texttt{sin}/\texttt{cos}/\texttt{sigmoid} pass) and pre-contracts the
mass integral into a $(N_{\rm lt}, N_z)$ cache before the $\theta$ loop
starts.  Inner body is pure multiply-add.  See
\texttt{docs/sigma\_prj\_py\_vs\_cpp\_report.tex} for the same
structural pattern applied to the full-shear evaluator.

\paragraph{Mass-axis shift.}  Must be \texttt{ln\_mass\_shift\_ =
0.0} (see \S\ref{sec:conventions}).  \texttt{HMF\_t} does the
rescale internally; adding another shift here was a latent
double-shift bug pre-2026-05.

% ============================================================================
\section{\texttt{bsel}: analytic sigmoid closure}
\label{sec:bsel}

\modname{Path}\,\pyfile{y3\_buzzard/bsel.py}. Python finisher that
reads $(P_1, I_1, J)$ from the datablock, marginalises
$(b_{\rm zero}, b_\infty)$ over the EMG $\ltr$-prior, and writes
the two scalars $(\Bsmall, \Blarge)$ plus $b_{\rm eff}$ per bin.

\subsection{Closed form (Costanzi-2026 Eq.~6 marginalised)}

\begin{align}
  b_{\rm sel}(\theta \,|\, \lob, \ltr, \zob)
    &= b_{\rm zero}(\ltr)\, [1 - \sigma(\theta)] + b_\infty(\ltr)\, \sigma(\theta) \\
  \langle b_{\rm sel}\rangle_{\ltr}(\theta)
    &= \Bsmall(\lob, \zob)\, [1 - \sigma(\theta)] + \Blarge(\lob, \zob)\, \sigma(\theta)
\end{align}
with
\begin{equation}
  \Bsmall = \langle b_{\rm zero}\rangle_{\ltr}, \qquad
  \Blarge = \langle b_\infty  \rangle_{\ltr},
\end{equation}
averages taken over $w(\ltr) = P(\lob | \ltr, \zob) \cdot
\int dM\, (dn/dM)\, P_{\rm HOD}(\ltr | M, z) \cdot M$.

\begin{algorithm}[H]
\caption{\texttt{bsel.execute}}
\KwConfig{\texttt{lob[]}, \texttt{zob[]}, \texttt{ltr\_lo},
 \texttt{ltr\_hi\_factor} (default 3), \texttt{n\_ltr} (default 128),
 \texttt{plob\_ltr\_params\_npz}}
\KwData{\blk{b\_sel\_marg\_P1}, \blk{b\_sel\_marg\_I1}, \blk{b\_sel\_marg\_J},
 \blk{haloModel/bias}, \blk{mass\_function}, \blk{cluster\_mor}, \blk{distances}}
\KwResult{\blk{b\_sel\_marginalised/\{lob, zob, theta, vals, b\_eff, b\_small, b\_large\}}}

Load EMG coeffs from \texttt{plob\_ltr\_params.npz} fallback\;
Build shared $M$ grid $[10^{13}, 10^{15.5}] \hMsun$, 100 log-nodes\;
\For{\rm each $(\lob_i, \zob_j)$ in the wall}{
  $b_{\rm eff}(\lob_i, \zob_j) = \dfrac{\int dM\, n(M) P(\lob_i | M, z) M b(M,z)}
                                       {\int dM\, n(M) P(\lob_i | M, z) M}$\;
  GL nodes $\ltr \in [1, \text{ltr\_hi\_factor} \cdot \lob_i]$, \texttt{n\_ltr} points\;
  $\Delta_{\rm RND} = P_1 + b_{\rm eff} \cdot (I_1 + J)$\;
  $\delta_{\rm prj}(\ltr) = (\lob - \ltr)/\Delta_{\rm RND} - 1$\;
  $b_\infty(\ltr) = b_{\rm eff}(1 + 0.13 \delta_{\rm prj})$\;
  $b_{\rm zero}(\ltr) = [(\lob-\ltr) - P_1 - b_\infty I_1] / J$\;
  $w(\ltr) = W_{\rm GL} \cdot$ EMG prior $\cdot$ HOD prior\;
  $\Bsmall = \int w(\ltr) b_{\rm zero}(\ltr) d\ltr / \int w\,d\ltr$\;
  $\Blarge = \int w(\ltr) b_\infty(\ltr) d\ltr / \int w\,d\ltr$\;
}
\end{algorithm}

\paragraph{Critical convention.}  The $n_M$ argument to
\texttt{\_compute\_b\_eff} and \texttt{\_ltr\_weight} must be
$dn/dM$, not $dn/d\ln M$ (the helpers multiply by $M$ internally to
produce $dn/d\ln M$).  The \blk{mass\_function/dndlnmh} datablock
value is $dn/d\ln M$, so \modname{bsel.py} divides by $M_{\rm grid}$
after the read.  Dropping this division was the 2026-05-06 ``extra
$M$'' bug that halved $\Bsmall$ and pushed it negative.

% ============================================================================
\section{\texttt{red\_shear\_prj}: two-halo projection}
\label{sec:shear_prj}

\modname{Path}\,\pyfile{src/models/sigma\_prj\_t.hh} +
\pyfile{src/modules/sigma\_prj\_cpu/ShearPrjEvaluator.cc}. The
flagship of the Costanzi-2026 pipeline: co-computes
$\Sprj, \DSprj$, and the tangential shear $\gamma_t^{\rm prj}$ on a
wall-grid of $(\zob, \lob, R)$ points.

\subsection{Integrand}

\begin{equation}
  \Sprj(R \,|\, \lob, \zob)
    = \int dz\, dM\, d\theta\ \
      w_z(z)\, \frac{dV}{d\Omega dz}\, n(M, z)\,
      \big[1 + b(M, z)\, b_{\rm sel}(\theta)\, \xinl\big]\,
      \Sigma_{\rm mis}(R \,|\, M, z, R_{\rm mis} = \theta D_A(\zob))\,
      \mathbb 1[\theta > \theta_{\rm excl}(z)].
\end{equation}
The $[1 + \dots]$ term decomposes into the \texttt{rnd} piece
(the 1) and the \texttt{cl} piece (the $b\,b_{\rm sel}\,\xinl$);
both are accumulated separately so the downstream user can study
them in isolation.

\subsection{Algorithm}

\begin{algorithm}[H]
\caption{\texttt{ShearPrjEvaluator::evaluate}}
\KwConfig{\texttt{n\_per\_seg} (default 30), \texttt{n\_zring},
 \texttt{n\_zouter}, \texttt{n\_lnm}, \texttt{R\_max\_cMpch},
 \texttt{lob\_centers}, \texttt{theta\_breakpoints}}
\KwData{\blk{b\_sel\_marginalised/\{b\_small, b\_large\}}, \blk{haloModel/bias}, \blk{xi\_nl},
 \blk{mass\_function}, NFW miscentered tables (on disk)}
\KwResult{\blk{sigma\_prj/\{vals, rnd, cl\}}, \blk{dsigma\_prj/\{vals, rnd, cl\}},
 \blk{red\_shear\_prj/\{vals, rnd, cl\}}}

\For{\rm each unique $(\lob, \zob)$ on the wall}{
  \tcp*[l]{Build $\theta$-grid ONCE per $(\lob, \zob)$ slice}
  Collect $R$ values on this slice: $\{R_k\}$\;
  Breakpoint set:
  $\{\theta_{\rm lo},\, \theta_{\rm excl,o},\, \theta_R(R_k),\, \theta_\lambda,\, 2\theta_\lambda,\, \theta_{\max}\}$\;
  Sort, dedupe, drop outside $[\theta_{\rm lo}, \theta_{\max}]$\;
  Log-GL \texttt{n\_per\_seg} nodes per segment\;

  \tcp*[l]{Interpolate $\Bsmall, \Blarge$ in $\zob$}
  Linearly interp \blk{b\_sel\_marginalised/\{b\_small,b\_large\}} to exact $\zob$\;
  Pre-tabulate $b_{\rm sel}(\theta) = \Bsmall(1-\sigma) + \Blarge\,\sigma$\;

  \tcp*[l]{Pre-contract z, lnM to per-slice caches}
  \For{$iz = 1, \dots, N_z$}{
    Per-$z$ photo-$z$ gate + $\theta_{\rm excl}(z)$\;
    \For{$iM = 1, \dots, N_{\ln M}$}{
      Accumulate rnd and cl M-weights into per-slice arrays\;
    }
  }

  \tcp*[l]{Outer $\theta$ loop, inner $(z, R)$}
  \For{\rm each $\theta_{it}$ on the slice}{
    $\Sigma_{\rm mis}(R, \theta_{it} D_A, M)$ from table\;
    Accumulate rnd and cl contributions per $R$\;
  }

  \For{\rm each $R_k$ on the slice}{
    $\Sprj_{\rm rnd}(R_k)$, $\Sprj_{\rm cl}(R_k)$\;
    $\DSprj_{\rm rnd}(R_k)$, $\DSprj_{\rm cl}(R_k)$\;
    $g_t^{\rm prj}(R_k) = (\DSprj \cdot \mathrm{SCI}) / (1 - \Sprj \cdot \mathrm{SCI})$\;
  }
}
\end{algorithm}

\subsection{Numerical recipes}

\paragraph{Per-$R$ breakpoint $\theta$-grid.}
The integrand has three distinct features: the exclusion step at
$\theta_{\rm excl,o}$, the sigmoid transition near
$\theta_0 = \theta_\lambda/2$, and the $\Sigma_{\rm mis}$ peak at
$\theta_R(R) = R / D_A$.  A single GL panel over
$[\theta_{\rm lo}, 2\theta_\lambda]$ under-resolves all three.
Splitting the integration interval at the sorted breakpoint set
$\{\theta_{\rm excl,o}, \theta_R(R_k), \theta_\lambda, 2\theta_\lambda, \theta_{\max}\}$
reduces the residual by $\gtrsim 10\times$ at
$n_{\rm per\_seg} = 30$ vs a single-panel 120-node grid.

\paragraph{LoS-slab exclusion (per-$z$, not 3-D ball).}  Before the
refactor the exclusion was a 3-D ball mask $|\Delta r| < R_{\rm excl}$.
The correct 2-D-plus-LoS formulation is $\theta > \theta_{\rm excl}(z)$
with $\cos\theta_{\rm excl}(z) = (\chi_z^2 + \chi_o^2 - R_{\rm excl}^2)/(2\chi_z\chi_o)$.
Per-$z$ gating in the inner loop, cheap.

\paragraph{$z$-contracted per-slice weights.}  For each $(\lob, \zob)$
slice, the inner $z, M$ integrals are contracted into scalar
(rnd) and 1-D in $\theta$ (cl) caches.  The outer $\theta$ loop
collapses to multiply-adds over the cache plus the
$\Sigma_{\rm mis}$ table lookup.  Mirrors the \modname{b\_sel\_marg}
structure (\S\ref{sec:b_sel_marg}) and gives a $\sim 2\times$
wall-time win over the naive $(z, \theta, M)$ nest.

\paragraph{Analytic $b_{\rm sel}(\theta)$.}  The evaluator does
\emph{not} spline the legacy 32-node $b_{\rm sel}(\theta)$ table; it
evaluates the sigmoid analytically from the $\zob$-interpolated
$(\Bsmall, \Blarge)$ scalars produced by \modname{bsel}.  No ringing,
no nearest-$\zob$ tie-break bug.

\paragraph{Three integration backends, same recipe.}  The module
ships with three evaluators sharing the same integrand
implementation:
\begin{itemize}
  \setlength\itemsep{2pt}
  \item \modname{ShearPrjEvaluator} (default): fixed-GL log panels.
    \textbf{$\sim 250$\,ms / sample.}
  \item \modname{ShearPrjGsl}: GSL QAGP over $z$ for tight tolerance
    diagnostics.  Off by default.
  \item \modname{ShearPrjCuhre}: adaptive Cuhre on $(z, \ln M)$ inner.
    Off by default.  $\sim 30\times$ slower, used for convergence
    regression tests.
\end{itemize}

% ============================================================================
\section{Timing \& precision optimisation audit}
\label{sec:timing}

\subsection{Headline timing trajectory}

\begin{table}[H]
\centering
\small
\begin{tabular}{l r r r r r}
\toprule
Change & \modname{halo\_model} & \modname{sel\_function} & \modname{NumCts} &
         \modname{red\_shear\_prj} & Total \\
\midrule
Baseline               & 0.365\,s & 0.392\,s & 0.083\,s & 0.270\,s & \textbf{1.71\,s} \\
\texttt{compute\_lensing\_2h = F}   & 0.178\,s & 0.385\,s & 0.083\,s & 0.270\,s & 1.31\,s \\
\texttt{\_plob\_params} 1-D path    & 0.178\,s & 0.301\,s & 0.083\,s & 0.270\,s & 1.25\,s \\
\texttt{\_cdf\_lob} in-place        & 0.178\,s & 0.280\,s & 0.083\,s & 0.270\,s & 1.22\,s \\
\texttt{n\_lnm = 192}               & 0.154\,s & 0.225\,s & 0.117\,s & 0.247\,s & \textbf{1.15\,s} \\
\bottomrule
\end{tabular}
\caption{Per-module wall time on \modname{smoke\_full\_sel\_cp\_camb.ini},
login node, single thread.  Cumulative savings: $\sim 560$\,ms per
sample ($-33\%$).}
\end{table}

\subsection{\texttt{n\_lnm} convergence study}

Sweep of \modname{sel\_function/n\_lnm}; NumCountsSel refines Cuhre
harder when the $S_{\rm stack}$ grid is coarser, so the
\modname{sel\_function} savings beyond $n_{\ln M} = 192$ are
\emph{reclaimed} by Cuhre and the \emph{total} pipeline gets slower.

\begin{table}[H]
\centering
\small
\begin{tabular}{r r r r r r}
\toprule
$n_{\ln M}$ & \modname{sel\_function} & \modname{NumCts} & \modname{RS1h} & total &
              $\Delta N_c$ vs 256 \\
\midrule
256  & 0.384\,s & 0.083\,s & 0.097\,s & 1.36\,s & ref \\
\textbf{192}  & \textbf{0.224\,s} & \textbf{0.121\,s} & 0.085\,s & \textbf{1.15\,s} & $+0.01\%$ \\
128  & 0.148\,s & 0.273\,s & 0.080\,s & 1.22\,s & $+0.03\%$ \\
 96  & 0.113\,s & 0.211\,s & 0.076\,s & 1.13\,s & $+0.07\%$ \\
 64  & 0.072\,s & 0.207\,s & 0.077\,s & 1.11\,s & $\mathbf{-4.55\%}$ \rlap{\ \textcolor{harvardcrimson}{$\leftarrow$ GL resonance, avoid}} \\
 48  & 0.055\,s & 0.262\,s & 0.062\,s & 1.18\,s & $+0.33\%$ \\
\bottomrule
\end{tabular}
\caption{Whole-pipeline optimum is \textbf{$n_{\ln M} = 192$}, not
the single-module optimum.  $n_{\ln M} = 64$ hits a pathological GL
resonance that produces a $4.5\%$ error on $N_c$; avoid.}
\label{tab:nlnm-sweep}
\end{table}

\subsection{Debugging arc (C++\,vs\,Py)}

Six bugs were uncovered and audited out over the April--May 2026
sprint; documented in full in the companion
\modname{sigma\_prj\_py\_vs\_cpp\_report.tex}.  In brief order of
appearance and fix:

\begin{enumerate}
  \setlength\itemsep{2pt}
  \item \textbf{\modname{mf\_tinker} mass-axis}: \blk{m\_h} stored in
    $\Omega_m \hMsun$ units, not $\hMsun$.  Fixed by setting
    \texttt{ln\_mass\_shift\_ = 0.0} in all C++ evaluators that
    read through \texttt{HMF\_t} (\texttt{HMF\_t} does the rescale
    at construction; a second shift was double-counting).
  \item \textbf{Growth-factor normalisation}: \modname{halo\_model\_cosmosis}
    used un-normalised $D(z)$ in the bias $\nu$-scale.  Fixed by
    dividing out $D(0)$ explicitly.
  \item \textbf{\modname{cp\_camb} \texttt{nonu\_fallback}}: the stub
    copied the total-matter $P(k)$ into \blk{cdm\_baryon\_power\_lin}
    even when the real nonu emulator was available.  Fixed by
    plugging \texttt{linear\_nonu\_pk\_path}.
  \item \textbf{Python HMF $z$-evolution}:
    \texttt{richness\_selection.HMF} used Tinker 2008 $z{=}0$
    constants at every $z$; CosmoSIS \modname{mf\_tinker} applies Table 4
    $(1+z)$ scaling on $A, a, b$.  Fixed by enabling
    \texttt{z\_evolution=True} in the Python HMF.
  \item \textbf{Stale \modname{lob\_center}}: hardcoded bin-3 centre
    was $100$ (legacy SDSS) instead of $130$ (DES Y3).  Fixed in
    \texttt{sp\_detail::lob\_center} and made configurable via
    \texttt{lob\_centers} ini knob.
  \item \textbf{Spurious $M$ in \modname{bsel.py} prior}: passed
    $dn/d\ln M$ into a helper that multiplies by $M$ internally.
    Dragged $\Bsmall(\lob{=}25, \zob{=}0.425)$ from $+35$ to $+17$
    and pushed the small-$R$ \texttt{cl} piece of $\DSprj$ by
    $\sim 2\times$.  Fixed by dividing by $M$ on datablock read.
\end{enumerate}

After all six fixes: $\Sprj, \DSprj$ agree to $\leq 1.4\%$ across
$R \in [0.1, 30]\,\hMpc$ on both \texttt{rnd} and \texttt{cl};
$(P_1, I_1, J)$ agree to $\leq 1\%$; $(\Bsmall, \Blarge)$ agree to
$\sim 1\%$.

% ============================================================================
\section{Knob cheat-sheet}
\label{sec:knobs}

\begin{table}[H]
\centering
\small
\begin{tabular}{l l l p{0.40\linewidth}}
\toprule
Module & Knob & Value & Why \\
\midrule
\modname{cp\_camb}            & \texttt{linear\_nonu\_pk\_path} & v2c nonu emu & proper nonu P(k) \\
\modname{halo\_model}         & \texttt{compute\_lensing\_1h}   & T for Shear1hSel & publishes Sigma\_nfw \\
\modname{halo\_model}         & \texttt{compute\_lensing\_2h}   & \textbf{F} here          & saves 180--200\,ms \\
\modname{sel\_function}       & \texttt{n\_lnm}                 & \textbf{192}             & whole-pipeline optimum (\S\ref{sec:timing}) \\
\modname{sel\_function}       & \texttt{N\_q}                   & 32                      & GL bracket nodes in ltr \\
\modname{sel\_function}       & \texttt{n\_z\_shared}           & 64                      & shared z grid size \\
\modname{b\_sel\_marg}        & \texttt{n\_theta}               & 10 (split at $\theta_\lambda$) & sufficient with split \\
\modname{b\_sel\_marg}        & \texttt{n\_lt}                  & 60                      & already converged at 16 \\
\modname{bsel}                & \texttt{n\_ltr}                 & 128                     & matches Python ref \\
\modname{bsel}                & \texttt{ltr\_hi\_factor}        & 3.0                     & matches Python $[1, 3\lob]$ \\
\modname{red\_shear\_prj}     & \texttt{n\_per\_seg}            & 30                      & per-segment GL nodes \\
\modname{red\_shear\_prj}     & \texttt{R\_max\_cMpch}          & 30.0                    & upper $\theta$ boundary \\
\modname{red\_shear\_prj}     & \texttt{lob\_centers}           & \{25, 37.5, 52.5, 130\} & DES Y3 arithmetic centres \\
\bottomrule
\end{tabular}
\caption{Post-audit defaults that should be carried into any chain
run.  These are the values live in \modname{smoke\_full\_sel\_cp\_camb.ini}
as of 2026-05-07.}
\end{table}

% ============================================================================
\section{Companion documents}

\begin{itemize}
  \setlength\itemsep{1pt}
  \item \modname{docs/sigma\_prj\_py\_vs\_cpp\_report.tex}: the full
    C++\,vs\,Python audit with per-$R$ \texttt{rnd}/\texttt{cl}
    tables for $\Sprj$, $\DSprj$, and $b_{\rm sel,marg}(\theta)$.
  \item \modname{BUILD\_PERLMUTTER.md}: exact toolchain and ninja build
    recipe; required to reproduce the C++ \texttt{.so} artefacts.
  \item \modname{CLAUDE.md}: convention cheatsheet for future
    development.
\end{itemize}

\end{document}
